### 2019-08-12 Microscopy Data.Rmd
more functions

---
title: "Microscopy Data"
author: "Juan P Barraza"
date: "6/19/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(pixmap)
library(rtiff)
library(reshape2)
library(gplots)
library(dplyr)
library(gdata)
library(ggplot2)
library(cowplot)
library(magrittr)
library(tidyr)
library(purrr)

PA14wt<- c('3-13-19', '3-19-19', '4-17-19')
PA14mut<- c('4-24-19', '4-25-19', '5-8-19')
PA01wt<- c('3-26-19','4-18-19')
PA01mut<- c('4-23-19')
names3<-c('rep1','rep2','rep3')
names2<-c('rep1','rep2')
names1<-c('rep1')
figuresPath<-"/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Figures1"#this is to save the figures as I am making them

GetNumber<-function(idx){
  if (idx<10){
    idx = paste0(0,idx)
  }
  return (idx)
}
```

```{r thresholds}
MapThreshold<- function(listSamples, reps){
  mappedData<-map2(listSamples, reps, GetThresholds)
  boundData<-bind_rows(mappedData)
  blah <-deparse(substitute(listSamples))#this is to obtain the name of the list as a string
  boundData<-boundData %>% mutate(Strain = blah)#this inserts the name into a column, but it wont work in just one line =(
  boundData<- gather(boundData, "Color", "Value", -Position, -Sample, -TimePoint, -Strain)
  boundData<-boundData %>% mutate(Condition = recode(Position, "0" = "Sa Mono", "1" = "Pa Mono", "2" = "Pa:Sa 1:1", "3" = "Pa:Sa 10:1", "4" = "Pa:Sa 1:100", "5" = "Pa:Sa 1:10") )%>% mutate(Species = recode(Color, "Red" = "Sa", "Green" = "Pa"))
  boundData$Condition<-as.factor(boundData$Condition)
  boundData$Condition<-factor(boundData$Condition, levels = c("Sa Mono", "Pa Mono", "Pa:Sa 1:1", "Pa:Sa 1:10", "Pa:Sa 1:100","Pa:Sa 10:1"))
  boundData$Species<-as.factor(boundData$Species)
  boundData$Strain<-as.factor(boundData$Strain)
  return(boundData)
}

GetThresholds<-function(sample, samplename){
  folder<- paste0('/Users/jupabago/Documents/Whiteley/PROJECTS/Spatial paper Data/Microscopy Data/Other Data/',sample,'/threshold')
  fileList<-list.files(path = folder, full.names = TRUE)
  growthDf<-data.frame()
  for (position in 0:5){
    currentPosition<-fileList[grep(paste0("p_",position,"."),fileList)]
    growthDf<-rbind(growthDf,ReadThresholdFile(currentPosition, position))}
  growthDf<-growthDf %>% mutate(Sample = samplename)
return(growthDf)
}

ReadThresholdFile<-function(file, position ){#This function outputs total number of pixels in the csv file from matlab
  datos <-read.csv(file, header = FALSE)
  colnames(datos)<-c('Red', 'Green')
  datos<-datos %>% mutate(Position = position)
  datos$Position<-as.factor(datos$Position)
  datos<-datos %>% mutate(TimePoint = row_number()-1)
  return(datos)
}

wtThresholds<-MapThreshold(PA14wt, names3)
mutThresholds<-MapThreshold(PA14mut, names3)
Pa01Thresholds<-MapThreshold(PA01wt, names2)
Pa01mutThresholds<-MapThreshold(PA01mut, names1)

thresholdData<-rbind(wtThresholds,mutThresholds,Pa01Thresholds,Pa01mutThresholds)

#This is to get an idea of the thresholding levels
#this is for all samples of Pa 
ggplot()+
  geom_jitter(data=thresholdData %>% filter(Species=="Pa"),height = 0, width = .1, aes(x=TimePoint, y=Value, shape = Sample, color = Strain), size=3)+
  facet_grid(.~Condition)
#This is only for PA14 since they seem to be the most consistent
PA14thresholdData <-thresholdData%>%filter(Species=="Pa") %>% filter(Strain %in% c("PA14wt", "PA14mut")) %>% group_by(Position, TimePoint, Strain, Condition, Species)%>% summarise(MeanThreshold = mean(Value), SDThreshold = sd(Value))
PA14average<-PA14thresholdData %>% group_by(Position, Strain, Condition, Species)%>% summarise(MeanThreshold = mean(MeanThreshold), SDThreshold = sd(MeanThreshold))

ggplot()+
  geom_point(data=thresholdData %>% filter(Species=="Pa")%>% filter(Strain %in% c("PA14wt", "PA14mut")), aes(x=TimePoint, y=Value, shape = Sample, color = Strain), size=2)+
  geom_line(data=PA14thresholdData,aes(x = TimePoint, y = MeanThreshold, color =Strain), size = 1)+
  geom_errorbar(data=PA14thresholdData,aes(x = TimePoint, ymin = MeanThreshold-SDThreshold, ymax = MeanThreshold+SDThreshold, color = Strain), width=.5, size = .5)+
  geom_hline(data = PA14average, aes(yintercept = MeanThreshold, color = Strain), size = 3)+
  facet_grid(.~Condition)

#Now we do Staph
SathresholdData <-thresholdData%>%filter(Species=="Sa") %>% group_by(Position, TimePoint, Strain, Condition, Species)%>% summarise(MeanThreshold = mean(Value), SDThreshold = sd(Value))
Saaverage<-SathresholdData %>% group_by(Position, Strain, Condition, Species)%>% summarise(MeanThreshold = mean(MeanThreshold), SDThreshold = sd(MeanThreshold))

ggplot()+
  geom_point(data=thresholdData %>% filter(Species=="Sa"), aes(x=TimePoint, y=Value, shape = Sample, color = Strain), size=2)+
  geom_line(data=SathresholdData,aes(x = TimePoint, y = MeanThreshold, color =Strain), size = 1)+
  geom_errorbar(data=SathresholdData,aes(x = TimePoint, ymin = MeanThreshold-SDThreshold, ymax = MeanThreshold+SDThreshold, color = Strain), width=2, size = .1)+
  geom_hline(data = Saaverage, aes(yintercept = MeanThreshold, color = Strain), size = 3)+
  facet_grid(.~Condition)

```

```{r threshold weighted growth curves}
#These are the dataframes that we need to merge
thresholdResultsDf<-merge(cleanResults, thresholdData) %>% mutate(WeightedLogValue = Value*LogTotals) %>% mutate(WeightedValue = Value *Totals) %>% mutate(correctedWeightedLogValue = WeightedLogValue+5)

thresholdData$Condition<-as.factor(thresholdData$Condition)
 boundData$Condition<-as.factor(boundData$Condition)
  ggplot()+
    geom_point(data=thresholdResultsDf%>% filter(Species == "Pa"), aes(x=TimePoint, y=correctedWeightedLogValue,  shape = Sample), size=2)+
    geom_point(data=cleanResults%>% filter(Species == "Pa"), aes(x=TimePoint, y=LogTotals,  shape = Sample), size=2, color = "red")+
    facet_grid(Strain~Condition)

ggplot()+
    geom_point(data=thresholdResultsDf%>% filter(Species == "Sa"), aes(x=TimePoint, y=correctedWeightedLogValue,  shape = Sample), size=2)+
    geom_point(data=cleanResults%>% filter(Species == "Sa"), aes(x=TimePoint, y=LogTotals,  shape = Sample), size=2, color = "red")+
    facet_grid(Strain~Condition)

#Linear regressions
#First get a df with only Sa Mono Red Channel
SaMonoTrhesholdDf<-thresholdData%>% filter(Condition == "Sa Mono") %>% filter(Species == "Sa")

#Linear regression for Thresholds
lm(SaMonoTrhesholdDf$Value ~ SaMonoTrhesholdDf$TimePoint) 
plot(SaMonoTrhesholdDf$TimePoint, SaMonoTrhesholdDf$Value)
abline( 0.18154, 0.00819 )

#Linear regression for Weighted Growth Curves
SaMonoWeightedGrowthCurve<-thresholdResultsDf%>% filter(Condition == "Sa Mono") %>% filter(Species == "Sa")
lmWeighted<-lm(SaMonoWeightedGrowthCurve$WeightedLogValue ~ SaMonoWeightedGrowthCurve$TimePoint) 
plot(SaMonoWeightedGrowthCurve$TimePoint, SaMonoWeightedGrowthCurve$WeightedLogValue)
abline( 1.01962, 0.06641 )

#Non-linear model
ggplot()+
  geom_smooth(data=SaMonoTrhesholdDf, aes(x=TimePoint, y=Value), size=2)+
  geom_point(data=SaMonoTrhesholdDf, aes(x=TimePoint, y=Value), size=2)

ggplot()+
  geom_smooth(data=SaMonoTrhesholdDf, method = lm, aes(x=TimePoint, y=Value), size=2)+
  geom_point(data=SaMonoTrhesholdDf, aes(x=TimePoint, y=Value), size=2)

#Now do the same for Pa
Pa14wtMonoTrhesholdDf<-thresholdData%>% filter(Condition == "Pa Mono") %>% filter(Species == "Pa") %>% filter(Strain == "PA14wt")
Pa14mutMonoTrhesholdDf<-thresholdData%>% filter(Condition == "Pa Mono") %>% filter(Species == "Pa") %>% filter(Strain == "PA14mut")
#Linear regression for Thresholds PA14wt
lm(Pa14wtMonoTrhesholdDf$Value ~ Pa14wtMonoTrhesholdDf$TimePoint) 
plot(Pa14wtMonoTrhesholdDf$TimePoint, Pa14wtMonoTrhesholdDf$Value)
abline( 0.08775, -0.001566 )

lm(Pa14mutMonoTrhesholdDf$Value ~ Pa14mutMonoTrhesholdDf$TimePoint) 
plot(Pa14mutMonoTrhesholdDf$TimePoint, Pa14mutMonoTrhesholdDf$Value)
abline( 0.082315, -0.001234 )

#Non-linear model
#PA14wt
ggplot()+
  geom_smooth(data=Pa14wtMonoTrhesholdDf, method = lm, aes(x=TimePoint, y=Value), size=2)+
  geom_point(data=Pa14wtMonoTrhesholdDf, aes(x=TimePoint, y=Value), size=2)
ggplot()+
  geom_smooth(data=Pa14wtMonoTrhesholdDf, aes(x=TimePoint, y=Value), size=2)+
  geom_point(data=Pa14wtMonoTrhesholdDf, aes(x=TimePoint, y=Value), size=2)
#PA14mut
ggplot()+
  geom_smooth(data=Pa14mutMonoTrhesholdDf, method = lm, aes(x=TimePoint, y=Value), size=2)+
  geom_point(data=Pa14mutMonoTrhesholdDf, aes(x=TimePoint, y=Value), size=2)
ggplot()+
  geom_smooth(data=Pa14mutMonoTrhesholdDf, aes(x=TimePoint, y=Value), size=2)+
  geom_point(data=Pa14mutMonoTrhesholdDf, aes(x=TimePoint, y=Value), size=2)

```

```{r biomass functions}
MapResults<- function(listSamples, reps){
  blah <-deparse(substitute(listSamples))
  mappedData<-map2(listSamples, reps, GetResults)
  boundData<-bind_rows(mappedData)
  boundData<-boundData %>% mutate(Strain = blah)
  boundData<-gather(boundData,"Color","Value",-Slice,-Position,-Sample,-TimePoint,-Strain) %>% mutate(LogValue=ifelse(Value>0,log10(Value),0))
  boundData<-boundData %>% mutate(Condition = recode(Position, "0" = "Sa Mono", "1" = "Pa Mono", "2" = "Pa:Sa 1:1", "3" = "Pa:Sa 10:1", "4" = "Pa:Sa 1:100", "5" = "Pa:Sa 1:10" )) %>% mutate(Species = recode(Color, "Red" = "Sa", "Green" = "Pa")) %>% mutate(ZDistance = Slice *0.44)
  boundData$Condition<-as.factor(boundData$Condition)
  boundData$Condition<-factor(boundData$Condition, levels = c("Sa Mono", "Pa Mono", "Pa:Sa 1:1", "Pa:Sa 1:10", "Pa:Sa 1:100","Pa:Sa 10:1"))
  boundData$Species<-as.factor(boundData$Species)
  boundData$Strain<-as.factor(boundData$Strain)
  return(boundData)
  }

GetResults<-function(sample, samplename){
  folder<- paste0('/Users/jupabago/Documents/Whiteley/PROJECTS/Spatial paper Data/Microscopy Data/Other Data/',sample,'/results')
  fileList<-list.files(path = folder, full.names = TRUE)
  growthDf<-data.frame()
  for (position in 0:5){
    currentPosition<-fileList[grep(paste0("p",position,"."),fileList)]
    for (timepoint in 0:17){
      currentTimePoint<-currentPosition[grep(paste0("t",GetNumber(timepoint),"_"),currentPosition)]
      growthDf<-rbind(growthDf,ReadResultsFile(currentTimePoint, position, timepoint))}}
  growthDf<-growthDf %>% mutate(Sample = samplename)
return(growthDf)
}

ReadResultsFile<-function(file, position, timepoint){
  datos <-read.csv(file, header = FALSE)
  colnames(datos)<-c('Red', 'Green', 'Slice')
  datos<-datos %>% mutate(Position = position)
  datos<-datos %>% mutate(TimePoint = timepoint)
  return(datos)
}
```
#Biomass per slice.

Where is the majority of the biomass at any given time and How dos it grow over time.
It additionally helps identify how reproducible are the experiments
```{r biomass per slice}
#Run the biomass functions on each set of samples and combine them at the end 
wtResults<-MapResults(PA14wt, names3)
mutResults<-MapResults(PA14mut, names3)
Pa01Results<-MapResults(PA01wt, names2)
Pa01mutResults<-MapResults(PA01mut, names1)

ResultsData<-rbind(wtResults,mutResults,Pa01Results,Pa01mutResults)

#This function uses facets to show the abundance of cells according to height over time
GraphSlices<-function (data, bug, facete){
  #first which bug is being selected
  if(bug=="Sa"){data <-data %>% filter(Species == "Sa")%>% filter(!Condition == "Pa Mono");nombre<-"S. aureus"}
  if(bug=="Pa"){data<-data %>% filter(Species == "Pa")%>% filter(!Condition == "Sa Mono");nombre<-"P. aeruginosa"}
  #then what kind of facet can be used
  if(missing(facete)){facete=facet_grid(Condition~TimePoint);condicion <- "Strain";forma <- NULL}
  else{facete <-facet_wrap(.~TimePoint,ncol = 6 );condicion <- "Condition";forma<-"Strain"}
  ggplot()+
    geom_point(data=data, aes_string(x="ZDistance", y="Value", color = condicion, shape = forma),size=1)+
    scale_y_log10(name=expression(paste('Biomass from voxel counts' )), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    #labs(title =expression(paste(italic(nombre), " biomass value per slice in SCFM2" )),  x = "Distance from bottom (z-slice)")+
    labs(title =paste0(nombre, " biomass value per slice in SCFM2" ),  x = "Distance from bottom (z-slice)")+
    facete
}

ggsave("Sa z-slice condition-timepoint1.png", plot = GraphSlices(ResultsData, "Sa"), path = figuresPath)
ggsave("Sa z-slice timepoint.png", plot = GraphSlices(ResultsData, "Sa", "BLAH"), path = figuresPath)
ggsave("Pa z-slice condition-timepoint.png", plot = GraphSlices(ResultsData, "Pa"), path = figuresPath)
ggsave("Pa z-slice timepoint.png", plot = GraphSlices(ResultsData, "Pa", "BLAH"), path = figuresPath)

#This graph shows how the distribution of biomass changes over time per experiment.
GraphSlicesTime<-function(datos, bug, strain){
  if(bug=="Sa"){datos<-datos %>% filter(Species == "Sa")%>% filter(!Condition == "Pa Mono");nombre<-"S. aureus"}
  if(bug=="Pa"){datos<-datos %>% filter(Species == "Pa")%>% filter(!Condition == "Sa Mono");nombre<-"P. aeruginosa"}
  ggplot()+
    geom_point(data=datos %>% filter(Strain == strain) , aes(x=ZDistance, y=Value, color = TimePoint),size=1)+
    scale_y_log10(name=expression(paste('Biomass from voxel counts' )), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title =paste(bug, "biomass value per slice in SCFM2 using",strain  ),  x = "Distance from bottom (z-slice)")+
    facet_grid(Sample~Condition)
}

ggsave("Sa z-slice PA14wt grid.png", plot = GraphSlicesTime(ResultsData, "Sa", "PA14wt"), path = figuresPath)
ggsave("Sa z-slice PA14mut grid.png", plot = GraphSlicesTime(ResultsData, "Sa", "PA14mut"), path = figuresPath)
ggsave("Pa z-slice PA14wt grid.png", plot = GraphSlicesTime(ResultsData, "Pa", "PA14wt"), path = figuresPath)
ggsave("Pa z-slice PA14mut grid.png", plot = GraphSlicesTime(ResultsData, "Pa", "PA14mut"), path = figuresPath)

```

Data from S. aureus. 
Remember that all mono cultures of Sa are essentially the same: Same strain, same inoculum, same condition. So this is the only condition where you have several replicates.
```{r biomass Sa}
#Now that I have analyzed the individual slices, I can analyze each time point
cleanResults<-ResultsData %>% group_by(TimePoint,Sample,Strain,Condition,Species) %>% summarise(Totals = sum (Value))%>% mutate(LogTotals = ifelse(Totals>0, log10(Totals), 0))
cleanResultsStats<-cleanResults %>% group_by(TimePoint,Strain,Condition,Species) %>% summarise(MeanTotals = mean(Totals), MeanLogTotals = mean(LogTotals), SDTotals = sd(Totals),SDLogTotals = sd(LogTotals))

#This is just to check the difference between the mean log and the log of the mean, which are quite different as it turns out...
cleanResultsStats1<-cleanResultsStats %>% mutate(CheckLog = MeanLogTotals^10) %>% mutate(CheckLog2 = log10(MeanTotals)) %>% mutate(DiffLog = abs(MeanLogTotals-CheckLog2)) %>% mutate(DiffTotals = abs(MeanTotals - CheckLog))

GraphGrowthCurves<-function (dataRaw, dataStats, bug){
  if(bug=="Sa"){dataRaw<-dataRaw %>% filter(Species == "Sa")%>% filter(!Condition == "Pa Mono");
  dataStats<-dataStats %>% filter(Species == "Sa")%>% filter(!Condition == "Pa Mono");nombre<-"S. aureus"}
  if(bug=="Pa"){dataRaw<-dataRaw %>% filter(Species == "Pa")%>% filter(!Condition == "Sa Mono");
  dataStats<-dataStats %>% filter(Species == "Pa")%>% filter(!Condition == "Sa Mono");nombre<-"P. aeruginosa"}
  ggplot()+
    geom_point(data=dataRaw, aes(x=TimePoint, y=Totals),  color = "Black", size=2)+
    geom_point(data=dataStats, aes(x=TimePoint, y=MeanTotals),  color = "Red", size=2)+
    geom_line(data=dataStats, aes(x=TimePoint, y=MeanTotals),  color = "Red", size=2)+
    scale_y_log10(limits= c(1e4,1e7), name=expression(paste('Biomass from voxel counts' )), breaks = c(10^(seq(5,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title =paste(bug, " growth in mono in SCFM2" ),  x = "Time (hours)",shape = "Culture condition", color = "Antiobiotic/Control")+
    facet_grid(Strain~Condition)
}
GraphGrowthCurvesExp<-function (dataRaw,dataStats, bug){
  if(bug=="Sa"){dataRaw<-dataRaw %>% filter(Species == "Sa")%>% filter(!Condition == "Pa Mono");
  dataStats<-dataStats %>% filter(Species == "Sa")%>% filter(!Condition == "Pa Mono");colore<-"Red"}
  if(bug=="Pa"){dataRaw<-dataRaw %>% filter(Species == "Pa")%>% filter(!Condition == "Sa Mono");
  dataStats<-dataStats %>% filter(Species == "Pa")%>% filter(!Condition == "Sa Mono");colore<-"Green"}
  ggplot()+
    geom_point(data=dataRaw, aes(x=TimePoint, y=LogTotals),  color = "Black", size=2)+
    geom_point(data=dataStats, aes(x=TimePoint, y=MeanLogTotals),  color = colore, size=2)+
    geom_line(data=dataStats, aes(x=TimePoint, y=MeanLogTotals),  color = colore, size=2)+
    labs(title =paste(bug, " growth in SCFM2" ),  x = "Time (hours)",shape = "Culture condition", color = "Antiobiotic/Control")+
    facet_grid(Strain~Condition)
}
GraphGrowthCurvesExpClean<-function (dataRaw, dataStats, bug){
  if(bug=="Sa"){dataStats<-dataStats %>% filter(Species == "Sa")%>% filter(!Condition == "Pa Mono");colore<-"Red"}
  if(bug=="Pa"){dataStats<-dataStats %>% filter(Species == "Pa")%>% filter(!Condition == "Sa Mono");colore<-"Green"}
  ggplot()+
    geom_point(data=dataStats, aes(x=TimePoint, y=MeanLogTotals,  color = Condition), size=2)+
    geom_line(data=dataStats, aes(x=TimePoint, y=MeanLogTotals,  color = Condition), size=2)+
    labs(title =paste(bug, " growth in SCFM2" ),  x = "Time (hours)",shape = "Culture condition", color = "Antiobiotic/Control")+
    facet_grid(Strain~.)
}

GraphGrowthCurvesExp(cleanResults,cleanResultsStats, "Sa")
GraphGrowthCurvesExp(cleanResults,cleanResultsStats, "Pa")
GraphGrowthCurvesExpClean(cleanResults,cleanResultsStats, "Sa")
GraphGrowthCurvesExpClean(cleanResults,cleanResultsStats, "Pa")

#First Isolate Sa Mono

cleanResultsSaMono<- cleanResultsSa %>% filter(Condition == "Sa Mono")

ggplot()+
  geom_point(data=cleanResultsSa %>% filter(Condition == "Sa Mono"), aes(x=TimePoint, y=Totals),  color = "Black", size=2)+
  geom_smooth(data=cleanResultsSaMono, aes(x=TimePoint, y=Totals), size=3, color = "red")+
  scale_y_log10(limits= c(1e4,1e7), name=expression(paste('Biomass from voxel counts' )), breaks = c(10^(seq(5,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title =expression(paste(italic("S. aureus"), " growth in mono in SCFM2" )),  x = "Time (hours)",shape = "Culture condition", color = "Antiobiotic/Control")

ggplot()+
  geom_point(data=cleanResultsSa, aes(x=TimePoint, y=Totals, shape = NULL,  color =Strain), size=3)+
  scale_y_log10(limits= c(1e4,1e7), name=expression(paste('Biomass' )), breaks = c(10^(seq(5,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(~Condition)+
  labs(title =expression(paste(italic("S. aureus"), " growth in mono or coculture in SCFM2" )),  x = "Time (hours)",shape = "Culture condition", color = "Antiobiotic/Control")

#Now Sa Co-culture
cleanResultsSaCo<- cleanResultsSa %>% filter(!Condition == "Sa Mono")
ggplot()+
  geom_point(data=cleanResultsSaCo, aes(x=TimePoint, y=Totals,  color = Strain), size=2)+
  geom_smooth(data=cleanResultsSaCo, aes(x=TimePoint, y=Totals, color = Strain), size=3)+
  scale_y_log10(limits= c(1e4,1e7), name=expression(paste('Biomass from voxel counts' )), breaks = c(10^(seq(5,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title =expression(paste(italic("S. aureus"), " growth in mono in SCFM2" )),  x = "Time (hours)",shape = "Culture condition", color = "Antiobiotic/Control")+
  facet_grid(~Condition)

```

```{r growth rates Sa}

#first try to fit it to a model fro exponential growth:
exponential.model <- lm(log(cleanResultsSaMono$Totals)~ cleanResultsSaMono$TimePoint)
summary(exponential.model)

#not very good... So I found the package "growthrates", so I'll try that
#https://cran.r-project.org/web/packages/growthrates/vignettes/Introduction.html
splitted.data <- multisplit(cleanResultsSaMono, c("Strain", "Sample"))
dat<-splitted.data[[1]]
fit <- fit_easylinear(dat$TimePoint, dat$Totals)
summary(fit)#summary of values
coef(fit) #exponential growth parameters
rsquared(fit)  # coefficient of determination (of log-transformed data)
deviance(fit)  # residual sum of squares of log-transformed data

par(mfrow = c(1, 2))
plot(fit, log = "y")
plot(fit)

many_fits <- all_easylinear(Totals ~ TimePoint | Strain + Sample, data = cleanResultsSaMono, h = 4)#the h is the number of points considered for fit

par(mfrow = c(3, 3))#establish the grid
plot(many_fits,log = "y")#plot many things
testdf<-data.frame(as.list(rsquared))#convert to data frame

#this is the mean rsquared for the growth curves. Not super useful, but it is the syntax for more interesting stuff
pa14GrowthRates<- testdf %>% select(contains("PA14")) %>% gather() %>% summarise(mean = mean(value))
#Now with all the conditions and find different growth rates

```

Data from P. aeruginosa

This is a completely different story since I used different Strains of Pa throughout the project. So here they are:
```{r biomass Pa}
cleanResultsPaMono<- cleanResultsPa %>% filter(Condition == "Pa Mono")
ggplot()+
  geom_point(data=cleanResultsPaMono, aes(x=TimePoint, y=Totals, color =Strain), size=2)+
  geom_smooth(data=cleanResultsPaMono, aes(x=TimePoint, y=Totals, color = Strain), size=3)+
  #geom_smooth(data=cleanResultsPaMono, method = "gam",  aes(x=TimePoint, y=Totals, color = Strain), linetype = "dashed", size=3)+
  scale_y_log10(limits= c(1e4,1e7), name=expression(paste('CFUs per mL' )), breaks = c(10^(seq(5,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))
  #facet_grid(~Strain)

#model to Fit into an exponential 
```

```{r biomass Pa and Sa}
ggplot()+
  geom_point(data=cleanResultsSaMono, aes(x=TimePoint, y=Totals),  color = "Black", size=2)+
  geom_point(data=cleanResultsPaMono, aes(x=TimePoint, y=Totals, color =Strain), size=2)+
  geom_smooth(data=cleanResultsPaMono, aes(x=TimePoint, y=Totals, color = Strain), size=3)+
  geom_smooth(data=cleanResultsSaMono, aes(x=TimePoint, y=Totals), size=3, color = "red")+
  #geom_smooth(data=cleanResultsPaMono, method = "gam",  aes(x=TimePoint, y=Totals, color = Strain), linetype = "dashed", size=3)+
  scale_y_log10(limits= c(1e4,1e7), name=expression(paste('Biomass in voxels' )), breaks = c(10^(seq(5,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title =expression(paste(italic("S. aureus"), " and ", italic("P. aeruginosa"), "growth in mono in SCFM2" )),  x = "Time (hours)")

ggplot()+
  scale_y_log10(limits= c(1e4,1e7), name=expression(paste('Biomass from voxel counts' )), breaks = c(10^(seq(5,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title =expression(paste(italic("S. aureus"), " growth in mono in SCFM2" )),  x = "Time (hours)",shape = "Culture condition", color = "Antiobiotic/Control")

```
## Including Plots

```{r aggregates}

MapAggregates<- function(listSamples, reps){
  mappedData<-map2(listSamples, reps, GetAggregates)
  boundData<-bind_rows(mappedData)
  blah <-deparse(substitute(listSamples))
  boundData<-boundData %>% mutate(Strain = blah)#needs to be two lines, it doesnt work if you do it directly
  return(boundData)
  }

GetAggregates<-function(sample, samplename){
  folder<- paste0('/Users/jupabago/Documents/Whiteley/PROJECTS/Spatial paper Data/Microscopy Data/Other Data/',sample,'/aggregateSizes')
  fileList<-list.files(path = folder, full.names = TRUE)
  growthDf<-data.frame()
  for (position in 0:5){
    currentPosition<-fileList[grep(paste0("p_",position,"."),fileList)]
    for (timepoint in 0:17){
      currentTimePoint<-currentPosition[grep(paste0("t_",GetNumber(timepoint),"p_"),currentPosition)]
      growthDf<-rbind(growthDf,ReadAggregateFile(currentTimePoint, position, timepoint))
      }
  }
  growthDf<-growthDf %>% mutate(Sample = samplename)
return(growthDf)
}

ReadAggregateFile<-function(file, position, timepoint){#This function outputs total number of pixels in the csv file from matlab
  datos <-read.csv(file, header = TRUE)
  colnames(datos)<-c('Size', 'Species')
  datos<-datos %>% mutate(Position = position)
  datos<-datos %>% mutate(TimePoint = timepoint)
  return(datos)
}

wtAggregates<-MapAggregates(PA14wt, names3)
mutAggregates<-MapAggregates(PA14mut, names3)
Pa01Results<-MapAggregates(PA01wt, names2)
Pa01mutResults<-MapAggregates(PA01mut, 'rep1')

```
You can also embed plots, for example:

```{r support functions, echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.