### 2019-08-14 CFU Data.Rmd
new file with the CFU data with the new SCFM2

---
title: "Growth Curves Paper"
author: "Juan P Barraza"
date: "8/14/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(pixmap)
library(rtiff)
library(reshape2)
library(gplots)
library(dplyr)
library(gdata)
library(ggplot2)
library(cowplot)
library(magrittr)
library(tidyr)
library(plotrix)
library(purrr)

#Data gathering functions:
GetGrowthCurveReplicate<- function (sample){
  bugCounts<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/Spatial paper Data/GrowthCurves/", sample, ".csv"),header = TRUE)
  bugCounts<- na.omit(bugCounts)
  bugCounts<-bugCounts %>% mutate(logCFU = log10(CFU))
  bugCounts$sample<-sample
  return(bugCounts)
}
GetGrowthCurveCondition<-function(datalist){
  mappedData<-map(datalist, GetGrowthCurveReplicate)#this line reads all the files on the list
  boundData<-bind_rows(mappedData)
  colnames(boundData)[5] <- "strain"
  colnames(boundData)[8] <- "condition"
  return(boundData)
}
#Graphing functions
##this one graphs the log values
GraphGrowthCurveLog<-function (rawData, statsData){
    ggplot()+
    geom_point(data=rawData, aes(x=time, y=logCFU,color = bug, shape = strain),   size=2)+
    geom_point(data=statsData, aes(x=time, y=logCFUmean,  color = bug, shape = strain), size=2)+
    geom_line(data=statsData, aes(x=time, y=logCFUmean,  color = bug, linetype = strain), size=2)+
    #geom_errorbar(data=statsData, aes(x = time, ymin = logCFUmean-logCFUsd, ymax = logCFUmean+logCFUsd, color = bug, linetype = strain), width = 0, size = 1)+
    geom_errorbar(data=statsData, aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse, color = bug, linetype = strain), width = 0, size = 1)+
    labs(title =paste("Growth in SCFM2" ),  x = "Time (hours)",shape = "Culture condition", color = "Bug")+
    facet_grid(.~condition)
}
##this one graphs the normal values
GraphGrowthCurve<-function (rawData, statsData){
    ggplot()+
    geom_point(data=rawData, aes(x=time, y=CFU,color = bug, shape = strain),   size=2)+
    geom_point(data=statsData, aes(x=time, y=CFUmean,  color = bug, shape = strain), size=2)+
    geom_line(data=statsData, aes(x=time, y=CFUmean,  color = bug, linetype = strain), size=2)+
    geom_errorbar(data=statsData, aes(x = time, ymin = CFUmean-CFUsd, ymax = CFUmean+CFUsd, color = bug, linetype = strain), width = 0, size = 1)+
    labs(title =paste("Growth in SCFM2" ),  x = "Time (hours)",shape = "strain", color = "Bug")+
    scale_y_log10(name=expression(paste('CFUs per mL' )), breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    facet_grid(.~condition)
}
```

```{r results}
#these are the list with the spreadsheets of experiments under different conditions
PA14 = c('2-11-19','2-7-19','2-5-19')
#get data
PA14Data<-GetGrowthCurveCondition(PA14)
#do stats
PA14DataStats<-PA14Data%>% group_by(condition, bug, time,strain) %>% summarise(CFUmean = mean (CFU), CFUsd = sd(CFU), CFUse = std.error(CFU), logCFUmean = mean (logCFU), logCFUsd = sd(logCFU),logCFUse = std.error(logCFU))
#Graph logs
GraphGrowthCurveLog(PA14Data,PA14DataStats)
#Graph regular numbers
GraphGrowthCurve(PA14Data,PA14DataStats)

```