### 2019-08-16 CFU Data.Rmd
added comparison of CFUs to Confocal biomass

---
title: "Growth Curves Paper"
author: "Juan P Barraza"
date: "8/14/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(pixmap)
library(rtiff)
library(reshape2)
library(gplots)
library(dplyr)
library(gdata)
library(ggplot2)
library(cowplot)
library(magrittr)
library(tidyr)
library(plotrix)
library(purrr)
library(moderndive)

#Data gathering functions:
GetGrowthCurveReplicate<- function (sample){
  bugCounts<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/Spatial paper Data/GrowthCurves/", sample, ".csv"),header = TRUE)
  bugCounts<- na.omit(bugCounts)
  bugCounts<-bugCounts %>% mutate(logCFU = log10(CFU))
  bugCounts$sample<-sample
  return(bugCounts)
}
GetGrowthCurveCondition<-function(datalist){
  mappedData<-map(datalist, GetGrowthCurveReplicate)#this line reads all the files on the list
  boundData<-bind_rows(mappedData)
  colnames(boundData)[5] <- "strain"
  colnames(boundData)[8] <- "condition"
  return(boundData)
}
#Graphing functions
##this one graphs the log values
GraphGrowthCurveLog<-function (rawData, statsData){
    ggplot()+
    geom_point(data=rawData, aes(x=time, y=logCFU,color = bug, shape = strain),   size=2)+
    geom_point(data=statsData, aes(x=time, y=logCFUmean,  color = bug, shape = strain), size=2)+
    geom_line(data=statsData, aes(x=time, y=logCFUmean,  color = bug, linetype = strain), size=2)+
    #geom_errorbar(data=statsData, aes(x = time, ymin = logCFUmean-logCFUsd, ymax = logCFUmean+logCFUsd, color = bug, linetype = strain), width = 0, size = 1)+
    geom_errorbar(data=statsData, aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse, color = bug, linetype = strain), width = 0, size = 1)+
    labs(title =paste("Growth in SCFM2" ),  x = "Time (hours)",shape = "Culture condition", color = "Bug")+
    facet_grid(.~condition)
}
##this one graphs the normal values
GraphGrowthCurve<-function (rawData, statsData){
    ggplot()+
    geom_point(data=rawData, aes(x=time, y=CFU,color = bug, shape = strain),   size=2)+
    geom_point(data=statsData, aes(x=time, y=CFUmean,  color = bug, shape = strain), size=2)+
    geom_line(data=statsData, aes(x=time, y=CFUmean,  color = bug, linetype = strain), size=2)+
    geom_errorbar(data=statsData, aes(x = time, ymin = CFUmean-CFUsd, ymax = CFUmean+CFUsd, color = bug, linetype = strain), width = 0, size = 1)+
    labs(title =paste("Growth in SCFM2" ),  x = "Time (hours)",shape = "strain", color = "Bug")+
    scale_y_log10(name=expression(paste('CFUs per mL' )), breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    facet_grid(.~condition)
}
```

```{r CFU results}
#these are the list with the spreadsheets of experiments under different conditions
PA14 = c('2-11-19','2-7-19','2-5-19')
#get data
PA14Data<-GetGrowthCurveCondition(PA14)
#do stats
PA14DataStats<-PA14Data%>% group_by(condition, bug, time,strain) %>% summarise(CFUmean = mean (CFU), CFUsd = sd(CFU), CFUse = std.error(CFU), logCFUmean = mean (logCFU), logCFUsd = sd(logCFU),logCFUse = std.error(logCFU))
#Graph logs
GraphGrowthCurveLog(PA14Data,PA14DataStats)
#Graph regular numbers
GraphGrowthCurve(PA14Data,PA14DataStats)

```

```{r Confocal vs CFUs data wrangling}
#first pull confocal data:
##functions to read data
MapResults<- function(listSamples, reps){
  blah <-deparse(substitute(listSamples))
  mappedData<-map2(listSamples, reps, GetResults)
  boundData<-bind_rows(mappedData)
  boundData<-boundData %>% mutate(Strain = blah)
  boundData<-gather(boundData,"Color","Value",-Slice,-Position,-Sample,-TimePoint,-Strain) %>% mutate(LogValue=ifelse(Value>0,log10(Value),0))
  boundData<-boundData %>% mutate(Condition = recode(Position, "0" = "Sa Mono", "1" = "Pa Mono", "2" = "Pa:Sa 1:1", "3" = "Pa:Sa 10:1", "4" = "Pa:Sa 1:100", "5" = "Pa:Sa 1:10" )) %>% mutate(Species = recode(Color, "Red" = "Sa", "Green" = "Pa")) %>% mutate(ZDistance = Slice *0.44)
  boundData$Condition<-as.factor(boundData$Condition)
  boundData$Condition<-factor(boundData$Condition, levels = c("Sa Mono", "Pa Mono", "Pa:Sa 1:1", "Pa:Sa 1:10", "Pa:Sa 1:100","Pa:Sa 10:1"))
  boundData$Species<-as.factor(boundData$Species)
  boundData$Strain<-as.factor(boundData$Strain)
  return(boundData)
  }
ReadResultsFile<-function(file, position, timepoint){
  datos <-read.csv(file, header = FALSE)
  colnames(datos)<-c('Red', 'Green', 'Slice')
  datos<-datos %>% mutate(Position = position)
  datos<-datos %>% mutate(TimePoint = timepoint)
  return(datos)
}
GetResults<-function(sample, samplename){
  folder<- paste0('/Users/jupabago/Documents/Whiteley/PROJECTS/Spatial paper Data/Microscopy Data/Other Data1/',sample,'/results')#the path needs to be "Other Data1"
  fileList<-list.files(path = folder, full.names = TRUE)
  growthDf<-data.frame()
  for (position in 0:5){
    currentPosition<-fileList[grep(paste0("p",position,"."),fileList)]
    for (timepoint in 0:17){
      currentTimePoint<-currentPosition[grep(paste0("t",GetNumber(timepoint),"_"),currentPosition)]
      growthDf<-rbind(growthDf,ReadResultsFile(currentTimePoint, position, timepoint))}}
  growthDf<-growthDf %>% mutate(Sample = samplename)
return(growthDf)
}

PA14wt<- c('3-13-19', '3-19-19', '4-17-19')#sample names 
PA14mut<- c('4-24-19', '4-25-19', '5-8-19')
names3<-c('rep1','rep2','rep3')

wtConfocalResults<-MapResults(PA14wt, names3)
mutConfocalResults<-MapResults(PA14mut, names3)
#combine them:
allConfocalResults<-rbind(wtConfocalResults,mutConfocalResults)
#Get stats:
##this adds up all the data on each slice and then gets the log of that total
allConfocalResults<-allConfocalResults %>% group_by(TimePoint,Sample,Strain,Condition,Species) %>% summarise(Totals = sum (Value))%>% mutate(LogTotals = ifelse(Totals>0, log10(Totals), 0))

##this calculates mean and error
allConfocalResultsStats<-allConfocalResults %>% group_by(TimePoint,Strain,Condition,Species) %>% summarise(MeanTotals = mean(Totals), SDTotals = sd(Totals), MeanLogTotals = mean(LogTotals), SDLogTotals = sd(LogTotals))

##remove all the conditions I didn't do in the plate counts and fix time points:
DataComparisonScope<-allConfocalResultsStats %>% filter (Strain %in% c("PA14wt","PA14mut")) %>% filter(Condition == "Pa:Sa 1:1") %>%  filter(TimePoint %in% c(0,3,7,11,15)) %>% select(Condition,TimePoint, Strain, Species, MeanTotals, MeanLogTotals) %>% mutate(time = ifelse(TimePoint>0, TimePoint+1, 0)) #The timePoint thing is because I start imaging after one hour and I don't have a time zero on scope

##remove conditions I didn't do in the scope and fix strain names:
DataComparisonCFU<-PA14DataStats %>% filter (condition == "Static") %>% select(bug, strain, CFUmean, logCFUmean, condition, time)%>% as.data.frame() %>% mutate(Strain = recode(DataComparisonCFU$strain, wt = "PA14wt", mut = "PA14mut"))

#name columns so that they match with the columns in the other df
colnames(DataComparisonCFU)<- c("Species", "strain", "CFUs", "logCFUs","Static", "time", "Strain")

#merge data frames and fix numeric for time
comparisonDf <- merge(DataComparisonCFU,DataComparisonScope,by=c("Species","Strain", "time"))
comparisonDf$time<-as.numeric(as.character(comparisonDf$time))

##Adding more data from the TOB experiments!
#first select relevant conditions and fix column names
DataComparisonCFUSa<-newSCFMDataSa%>%  filter(!culture == "mono") %>% filter(antibiotic == "Control") %>% select(CFU, culture, bug, timePoint, antibiotic, logTranformed,sample, condition) %>% group_by(culture, timePoint) %>% summarise(meanCFU = mean(CFU), meanLogCFU = mean(logTranformed))  
DataComparisonCFUSa<-DataComparisonCFUSa%>% as.data.frame() %>%mutate(Strain = recode(DataComparisonCFUSa$culture, wt = "PA14wt", mut = "PA14mut"))

colnames(DataComparisonCFUSa)<-c("strain", "time", "CFUs", "logCFUs", "Strain")
#then do the same for the scope data
DataComparisonScopeSa<-allConfocalResultsStats %>% filter(Species == "Sa")%>% filter (Strain %in% c("PA14wt","PA14mut")) %>% filter(Condition == "Pa:Sa 1:1") %>%  filter(TimePoint %in% c(0,2,3,4)) %>% select(Condition,TimePoint, strain, Species, MeanTotals, MeanLogTotals) %>% mutate(time = ifelse(TimePoint>0, TimePoint+1, 0)) #The timePoint thing is because I start imaging after one hour and I don't have a time zero on scope
#merge them!
comparisonSaDf <- merge(DataComparisonCFUSa,DataComparisonScopeSa,by=c("time","Strain"))

```

```{r graph CFU vs Confocal}
#Graph both Pa and Sa
##Original data as input:
ggplot()+
    geom_point(data=comparisonDf, aes(x=CFUs, y=MeanTotals,color = Species, shape = Strain),   size=4)+
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))

##Log-tranformed data as input:
ggplot()+
    geom_point(data=comparisonDf, aes(x=logCFUs, y=MeanLogTotals, color = Species, shape = Strain),   size=4)

#Pa seems to be pretty straightforward, lets do linear regression:
Pa14LinearModel <- lm(MeanLogTotals ~logCFUs, data = comparisonDf %>% filter (Species == "Pa"))
coefficients(Pa14LinearModel)
get_regression_summaries(Pa14LinearModel, print = TRUE)

##Log-tranformed data as input for PA14 only:
ggplot()+
  geom_abline(intercept = 1.9623899, slope = 0.4639804, color = "red", size = 3 )+
  geom_point(data=comparisonDf%>% filter(Species == "Pa"), aes(x=logCFUs,y=MeanLogTotals,  color = Species, shape = Strain),   size=4)
#then fit it without the zero point since that was not even at the same time
Pa14LinearModel1 <- lm(MeanLogTotals ~logCFUs, data = comparisonDf %>% filter (Species == "Pa") %>% filter(!time == 0))
get_regression_summaries(Pa14LinearModel1, print = TRUE)
  ggplot()+
    geom_abline(intercept =-2.882887,slope = 1.006844, color = "red", size = 1, linetype = "dashed" )+
    geom_point(data=comparisonDf %>% filter(Species == "Pa"), aes(x=logCFUs,y=MeanLogTotals,  color = time, shape = Strain),size=4)+
    scale_color_gradient(low="blue", high="red")+
    annotate("text", x = 7.5, y = 6, label = expression(paste("y=1.007x-2.883  ", R^2, "= 0.9")), color = "red", size = 8)+
    labs(title =expression(paste(italic("P. aeruginosa"), " CFU vs Confocal Biomass")), x = "Log CFUs",y = "Biomass from Scope")

#Now we go to Sa
SaLinearModel <- lm(MeanLogTotals ~logCFUs, data = comparisonDf %>% filter (Species == "Sa"))
coefficients(SaLinearModel)
get_regression_summaries(SaLinearModel, print = TRUE)

SaLinearModel1 <- lm(MeanLogTotals ~logCFUs, data = comparisonDf %>% filter (Species == "Sa")%>% filter(!time == 0))
coefficients(SaLinearModel1)
get_regression_summaries(SaLinearModel1, print = TRUE)

ggplot()+
  geom_point(data=comparisonDf %>% filter(Species == "Sa"), aes(x=logCFUs,y=MeanLogTotals,  color = time, shape = Strain),size=4)+
  geom_abline(intercept =25.73, slope = -2.44, color = "red", size = 1, linetype = "dashed" )+
  scale_color_gradient(low="blue", high="red")+
  annotate("text", x = 7.9, y = 4.75, label = expression(paste("y=-2.44x+25.73  ", R^2, "= 0.619")), color = "red", size = 8)+
  labs(title =expression(paste(italic("S. aureus"), " CFU vs Confocal Biomass")), x = "Log CFUs",y = "Biomass from Scope")

#Adding the early time points of Staph:
SaLinearModel2 <- lm(MeanLogTotals ~logCFUs, data = comparisonSaDf%>% filter(!time == 0))
coefficients(SaLinearModel2)
get_regression_summaries(SaLinearModel2, print = TRUE)
ggplot()+
  geom_point(data=comparisonSaDf %>% filter(Species == "Sa"), aes(x=logCFUs,y=MeanLogTotals,  color = time, shape = Strain),size=4)+
  geom_abline(intercept =-2.693, slope = 0.928, color = "red", size = 1, linetype = "dashed" )+
  scale_color_gradient(low="blue", high="red")+
  annotate("text", x = 7.9, y = 5.2, label = expression(paste("y=0.928x-2.693 ", R^2, "= 0.59")), color = "red", size = 8)+
  labs(title =expression(paste(italic("S. aureus"), " CFU vs Confocal Biomass")), x = "Log CFUs",y = "Biomass from Scope")

```

```{r using old Sa data instead? Nope, didnt work}
#Change the function so that it takes the old data
GetResults<-function(sample, samplename){
  folder<- paste0('/Users/jupabago/Documents/Whiteley/PROJECTS/Spatial paper Data/Microscopy Data/Other Data/',sample,'/results')#the path needs to be "Other Data"
  fileList<-list.files(path = folder, full.names = TRUE)
  growthDf<-data.frame()
  for (position in 0:5){
    currentPosition<-fileList[grep(paste0("p",position,"."),fileList)]
    for (timepoint in 0:17){
      currentTimePoint<-currentPosition[grep(paste0("t",GetNumber(timepoint),"_"),currentPosition)]
      growthDf<-rbind(growthDf,ReadResultsFile(currentTimePoint, position, timepoint))}}
  growthDf<-growthDf %>% mutate(Sample = samplename)
return(growthDf)
}
wtConfocalResults1<-MapResults(PA14wt, names3)
mutConfocalResults1<-MapResults(PA14mut, names3)
#combine them:
allConfocalResults1<-rbind(wtConfocalResults1,mutConfocalResults1)
#Get stats:
##this adds up all the data on each slice and then gets the log of that total
allConfocalResults1<-allConfocalResults1%>% group_by(TimePoint,Sample,Strain,Condition,Species) %>% summarise(Totals = sum (Value))%>% mutate(LogTotals = ifelse(Totals>0, log10(Totals), 0))

##this calculates mean and error
allConfocalResultsStats1<-allConfocalResults1%>% group_by(TimePoint,Strain,Condition,Species) %>% summarise(MeanTotals = mean(Totals), SDTotals = sd(Totals), MeanLogTotals = mean(LogTotals), SDLogTotals = sd(LogTotals))

##remove all the conditions I didn't do in the plate counts and fix time points:
DataComparisonScope1<-allConfocalResultsStats1 %>% filter (Strain %in% c("PA14wt","PA14mut")) %>% filter(Condition == "Pa:Sa 1:1") %>%  filter(TimePoint %in% c(0,3,7,11,15)) %>% select(Condition,TimePoint, Strain, Species, MeanTotals, MeanLogTotals) %>% mutate(time = ifelse(TimePoint>0, TimePoint+1, 0)) #The timePoint thing is because I start imaging after one hour and I don't have a time zero on scope

comparisonDf1 <- merge(DataComparisonCFU,DataComparisonScope1,by=c("Species","Strain", "time"))
comparisonDf1$time<-as.numeric(as.character(comparisonDf1$time))

ggplot()+
    geom_point(data=comparisonDf1 %>% filter(Species == "Sa"), aes(x=logCFUs, y=MeanLogTotals, color = time, shape = Strain),   size=4)+
  scale_color_gradient(low="blue", high="red")
```