### 2019-09-26 PropOc Data.Rmd

---
title: "Proportional Occupancy"
author: "Juan P Barraza"
date: "9/11/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(pixmap)
library(rtiff)
library(reshape2)
library(gplots)
library(dplyr)
library(gdata)
library(ggplot2)
library(cowplot)
library(magrittr)
library(tidyr)
library(purrr)
library(moderndive)
```

```{r read data, include=FALSE}
#names of samples
wtsamplesList<-c("PA14wt_1","PA14wt_2","PA14wt_3")
mutsamplesList<-c("PA14mut_1","PA14mut_2","PA14mut_3")

ReadPropOcFile<-function(sample, strain){#This function outputs total number of pixels in the csv file from matlab
  file<- paste0('/Users/jupabago/Documents/Whiteley/PROJECTS/Spatial paper Data/Microscopy Data/PropOc Data1/',sample,'.csv')
  datos <-read.csv(file, header = TRUE)
  datos<-datos %>% mutate(TimePoint = timePoint+1)
  datos<-datos %>% mutate(Sample = sample)
  datos$Sample<-as.factor(datos$Sample)
  datos<-datos %>% mutate(Strain = strain)
  datos$Strain<-as.factor(datos$Strain)
  datos<-datos %>% mutate(Relation = recode(source, "finalGG" = "Pa->Pa","finalRG" = "Sa->Pa","finalRR" = "Sa->Sa","finalGR" = "Pa->Sa"))%>%mutate(Condition = recode(position, "2" = "Pa:Sa 1:1", "3" = "Pa:Sa 10:1", "4" = "Pa:Sa 1:100", "5" = "Pa:Sa 1:10"))
  datos$Relation<-as.factor(datos$Relation)
  datos$Relation<-factor(datos$Relation, levels = c("Pa->Pa", "Sa->Sa", "Sa->Pa","Pa->Sa"))
  datos$Condition<-as.factor(datos$Condition)
  datos$Condition<-factor(datos$Condition, levels = c("Pa:Sa 1:1", "Pa:Sa 1:10", "Pa:Sa 1:100","Pa:Sa 10:1"))
  return(datos)
}
#combine the data from all samples
wtpropOcData.list<-map2(wtsamplesList, "wt", ReadPropOcFile)
mutpropOcData.list<-map2(mutsamplesList, "mut", ReadPropOcFile)
propOcData<-bind_rows(wtpropOcData.list,mutpropOcData.list)
```

```{r graphing all data}
#Start building some graphs:
graphAllData<-function(relation){
ggplot()+
  geom_point(data = propOcData%>% filter(Relation == relation), aes (x = distance, y = normPropOccup, color = Strain))+
  facet_grid(Condition~TimePoint)+
  ylab("Proportional occupancy")+
  ggtitle(paste0("Proportional occupancy ",relation, " (All data)"))
}
graphAllData("Pa->Pa")
graphAllData("Sa->Sa")
graphAllData("Sa->Pa")
graphAllData("Pa->Sa")
```

```{r graph stats}
#do stats:
propOcData.stats<-propOcData %>% group_by(distance, Relation, Condition, Strain, TimePoint) %>% summarize(MeanPropOc = mean(normPropOccup), SDPropOc = sd(normPropOccup)) 
#graph them:
graphStatsData<-function(relation){
ggplot()+
  geom_line(data = propOcData.stats%>% filter(Relation == relation), aes (x = distance, y = MeanPropOc, color = Strain))+
  #geom_errorbar(data = propOcData.stats%>% filter(Relation == relation), aes (x = distance, ymin = MeanPropOc-SDPropOc,ymax = MeanPropOc+SDPropOc, color = Strain))+
  facet_grid(Condition~TimePoint)+
  ylab("Proportional Occupancy")+
  ggtitle(paste0("Proportional occupancy ",relation, " (Stats)"))
}
graphStatsData("Pa->Pa")
graphStatsData("Sa->Sa")
graphStatsData("Sa->Pa")
graphStatsData("Pa->Sa")
```