### 2020-03-13 R Analysis Scripts/PropOc Data 3D.Rmd
New script analysis

---
title: "PropOcData2"
author: "Juan P Barraza"
date: "3/5/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(pixmap)
library(rtiff)
library(reshape2)
library(gplots)
library(gdata)
library(plyr)
library(dplyr)
library(ggplot2)
library(cowplot)
library(tidyr)
library(purrr)
library(plotrix)
library(ggsignif)
library(rdist)

source('/Users/jupabago/Documents/Whiteley/PROJECTS/Rpackage/PropOc Package/R/PropOcFunctions2.0.R') #using the new 3D script with capped spheres.
```

```{r Get Prop Oc Data}
GetNewPropOcOralData<-function(sample){
  imagesFolder<-"/Users/jupabago/Documents/Whiteley/PROJECTS/Oral Penn/Images/WT&Mut&GtfB/"
  folder<-paste0(imagesFolder, sample, "/combined Bin",sample) 
  fileList<-list.files(path = folder, full.names = TRUE)
  fileList2<-fileList[grep("tif",fileList)]
  propOc<- NormRawPipeline(SlicesTo3D(fileList2,"red"),SlicesTo3D(fileList2,"green"),xydimSearch = 54, zdimSearch = 9,sampleSize = 1000,xyRealDim = 0.156,zRealDim = 0.93, pipeBinSize = 1)
  return(propOc)
}

samples = c("WT1","WT2","WT3","Mut1","Mut2","Mut3","GtfB1","GtfB2","GtfB3")

propOcDataList<-map(samples, GetNewPropOcOralData)#get the data
propOcDataNames<-map2(propOcDataList, samples, ~.x %>% mutate(Sample = .y))#add sample column
propOcData<-bind_rows(propOcDataNames)#it's aliveeeee
write.csv(propOcData,"/Users/jupabago/545455Documents/Whiteley/PROJECTS/Oral Penn/Images/WT&Mut&GtfB/ PropOc Data AllData.csv")
```

```{r PropOc DATA}

GetPropOcData<-function(sample, timepoints, positions){
  folder<- paste0('/Volumes/Seagate Backup Plus Drive/Good images/',sample,'/images1')
  growthDf<-data.frame()#data frame to add all timepoints and positions
  fileList<-list.files(path = folder, full.names = TRUE)
  print(folder)
  for (position in 0:positions){
    currentPosition<-fileList[grep(paste0("p",position,"_"),fileList)]
    print(paste0('analyzing position ', position))
    for (timepoint in 0:timepoints){
      currentTimePoint<-currentPosition[grep(paste0("t",GetSlice(timepoint),"_"),currentPosition)]
      dataTimePoint<-NormRawPipeline(SlicesTo3D(currentTimePoint,"red"),SlicesTo3D(currentTimePoint,"green"), 
                                               #xydimSearch=114,zdimSearch = 70, sampleSize=1,xyRealDim=0.264,
                                     xydimSearch=11,zdimSearch = 7, sampleSize=100,xyRealDim=0.264,
                                               #zRealDim = 0.44, pipeBinSize = 1)
                                     zRealDim = 0.44, pipeBinSize = 1)
      dataTimePoint$TimePoint <- timepoint
      dataTimePoint$Position <- position
      growthDf<-rbind(growthDf,dataTimePoint)
      }
    }
  return(growthDf)
  }

GetSlice<-function(idx){
if (idx<10){num = paste0(0,idx)}
  else
    {num = idx}
  return(num)}

test1<-GetPropOcData("3-13-19", 0, 0)
test2<-GetPropOcData("3-13-19", 1, 1)

propOcPA14wt_1.1<-GetImageFiles("/media/whiteley/Seagate Backup Plus Drive/Good images/3-13-19/images1", timepoints = 10, positions = 5)
propOcPA14wt_2.1<-GetImageFiles("/media/whiteley/Seagate Backup Plus Drive/Good images/3-19-19/images1", timepoints = 10, positions = 5)
propOcPA14wt_3.1<-GetImageFiles("/media/whiteley/Seagate Backup Plus Drive/Good images/4-17-19/images1", timepoints = 10, positions = 5)

propOcPA14mut_1.1<-GetImageFiles("/media/whiteley/Seagate Backup Plus Drive/Good images/4-24-19/images1", timepoints = 10, positions = 5)
propOcPA14mut_2.1<-GetImageFiles("/media/whiteley/Seagate Backup Plus Drive/Good images/4-25-19/images1", timepoints = 10, positions = 5)
propOcPA14mut_3.1<-GetImageFiles("/media/whiteley/Seagate Backup Plus Drive/Good images/5-8-19/images1", timepoints = 10, positions = 5)

write.csv(propOcPA14wt_1.1, file = "/media/whiteley/Seagate Backup Plus Drive/Good images/PropOc Data/results1/PA14wt_1.csv", row.names = FALSE)
write.csv(propOcPA14wt_2.1, file = "/media/whiteley/Seagate Backup Plus Drive/Good images/PropOc Data/results1/PA14wt_2.csv", row.names = FALSE)
write.csv(propOcPA14wt_3.1, file = "/media/whiteley/Seagate Backup Plus Drive/Good images/PropOc Data/results1/PA14wt_3.csv", row.names = FALSE)

write.csv(propOcPA14mut_1.1, file = "/media/whiteley/Seagate Backup Plus Drive/Good images/PropOc Data/results1/PA14mut_1.csv", row.names = FALSE)
write.csv(propOcPA14mut_2.1, file = "/media/whiteley/Seagate Backup Plus Drive/Good images/PropOc Data/results1/PA14mut_2.csv", row.names = FALSE)
write.csv(propOcPA14mut_3.1, file = "/media/whiteley/Seagate Backup Plus Drive/Good images/PropOc Data/results1/PA14mut_3.csv", row.names = FALSE)

```