### 2020-04-22 R Analysis Scripts/PropOc Data 3D.Rmd
Adding whole data sets to the analysis instead of just the mean and SD

---
title: "PropOcData2"
author: "Juan P Barraza"
date: "3/5/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(pixmap)
library(rtiff)
library(reshape2)
library(gplots)
library(gdata)
library(plyr)
library(dplyr)
library(ggplot2)
library(cowplot)
library(tidyr)
library(purrr)
library(plotrix)
library(ggsignif)
library(rdist)
library("vcd")#this is fro fitting models to data
library(viridis)#color palette
library(gganimate)#gifs for ggplot

source('/Users/jupabago/Documents/Whiteley/PROJECTS/Rpackage/PropOc Package/R/PropOcFunctions2.0.R') 
#using the new 3D script with capped spheres.
```

```{r functions to get PropOc Data}
GetPropOcData<-function(sample, timepoints, positions){
  folder<- paste0('/Volumes/Seagate Backup Plus Drive/Good images/',sample,'/imagesSubtractRed')
  growthDf<-data.frame()#data frame to add all timepoints and positions
  fileList<-list.files(path = folder, full.names = TRUE)
  print(folder)
  for (position in 0:positions){
    currentPosition<-fileList[grep(paste0("p",position,"_"),fileList)]
    print(paste0('analyzing position ', position))
    for (timepoint in 0:timepoints){
      currentTimePoint<-currentPosition[grep(paste0("t",GetSlice(timepoint),"_"),currentPosition)]
      dataTimePoint<-NormRawPipeline(SlicesTo3D(currentTimePoint,"red"),SlicesTo3D(currentTimePoint,"green"), 
                                     xydimSearch=114,zdimSearch = 70, sampleSize=1000,xyRealDim=0.264,
                                     zRealDim = 0.44, pipeBinSize = 1)
      dataTimePoint$TimePoint <- timepoint
      dataTimePoint$Position <- position
      growthDf<-rbind(growthDf,dataTimePoint)
      }
    }
  return(growthDf)
  }

GetSlice<-function(idx){
if (idx<10){num = paste0(0,idx)}
  else
    {num = idx}
  return(num)}
```
```{r run functions}
wt1<-GetPropOcData("3-13-19",6,2)
wt2<-GetPropOcData("3-19-19",6,2)
wt3<-GetPropOcData("4-17-19",6,2)
mut1<-GetPropOcData("4-24-19",6,2)
mut2<-GetPropOcData("4-25-19",6,2)
mut3<-GetPropOcData("5-8-19",6,2)

```
```{r Read data files}

Get3DPropOcData<-function(strain, biologicalRep, experiment){
  propOcData1<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataRedSub1/", strain,biologicalRep,"-1000-",1,".csv"),header = T)
  propOcData1<-propOcData1 %>% mutate(Strain = strain,BioRep = biologicalRep, RunRep = 1)
  propOcData2<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataRedSub1/", strain,biologicalRep,"-1000-",2,".csv"),header = T)
  propOcData2<-propOcData2 %>% mutate(Strain = strain,BioRep = biologicalRep, RunRep = 2)
  propOcData3<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataRedSub1/", strain,biologicalRep,"-1000-",3,".csv"),header = T)
  propOcData3<-propOcData3 %>% mutate(Strain = strain,BioRep = biologicalRep, RunRep = 3)
  allData<-bind_rows(propOcData1,propOcData2,propOcData3)
  allData<-allData %>%  mutate(CV = SDPropOcc/MeanPropOcc) %>% mutate(Experiment = experiment)
  allData<-allData %>% 
    mutate(Relation = recode(source, "finalGG" = "Pa->Pa","finalRG" = "Sa->Pa","finalRR" = "Sa->Sa","finalGR" = "Pa->Sa")) %>%
    mutate(Condition = recode(Position, "2" = "Pa:Sa 1:1")) %>%
    mutate(Interaction = recode(source, "finalGG" = "Self","finalRG" = "Other","finalRR" = "Self","finalGR" = "Other")) %>%
    mutate(FocusSpecies = recode(source, "finalGG" = "Pa","finalRG" = "Sa","finalRR" = "Sa","finalGR" = "Pa")) %>%
    mutate(TargetSpecies = recode(source, "finalGG" = "Pa","finalRG" = "Pa","finalRR" = "Sa","finalGR" = "Sa")) %>% 
    mutate(Time = recode(TimePoint, "0" = "Hour 1","1" = "Hour 2","2" = "Hour 3","3" = "Hour 4","4" = "Hour 5","5" = "Hour 6"))

  allData<-allData %>% filter(TimePoint<6 & Distance<30)#last distance is always off for some reason =(
  }

wt1<-Get3DPropOcData("wt", 1, "wt1")
wt2<-Get3DPropOcData("wt", 2, "wt2")
wt3<-Get3DPropOcData("wt", 3, "wt3")
wt3<-wt3 %>% filter(!RunRep==2|!TimePoint==1)

mut1<-Get3DPropOcData("mut", 1, "mut1")
mut2<-Get3DPropOcData("mut", 2, "mut2")
mut3<-Get3DPropOcData("mut", 3, "mut3")

combinedData<-bind_rows(wt1,wt2,wt3,mut1,mut2,mut3)
combinedData<-combinedData %>% filter(RunRep==1)#I just saw that theyr are all the same, so no point on dragging that much data
```
```{r checking the replicate runs}
PlotPropOcAll <- function(df, name) {
  ggplot()+
    geom_point(data=df, aes(x=Distance,y=MeanPropOcc, color = Relation, shape = factor(RunRep)),size=4)+
    geom_errorbar(data=df, aes(x=Distance,ymin=MeanPropOcc-SDPropOcc,ymax=MeanPropOcc+SDPropOcc, color = Relation, linetype = factor(RunRep)),size=4)+
    facet_wrap(TimePoint~Relation, ncol = 4)+
    scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title =name)+
    theme_cowplot()+
    background_grid()
  }
PlotPropOcAll(wt1,"wt1")
PlotPropOcAll(wt2,"wt2")
PlotPropOcAll(wt3,"wt3")
PlotPropOcAll(mut1,"mut1 ")
PlotPropOcAll(mut2,"mut2")
PlotPropOcAll(mut3,"mut3")
```

```{r Plot comparison of prop oc between Strains}
ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Pa->Pa"),aes(x=Distance,y=MeanPropOcc, color = Strain, shape =factor(BioRep) ),size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap( ~ TimePoint, ncol = 3)+
  labs(title ="Pa->Pa")+
  theme_cowplot()

ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Sa->Pa"),aes(x=Distance,y=MeanPropOcc, color = Strain, shape =factor(BioRep) ),size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap( ~ TimePoint, ncol = 3)+
  labs(title ="Sa->Pa")+
  theme_cowplot()

ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Pa->Sa"),aes(x=Distance,y=MeanPropOcc, color = Strain, shape =factor(BioRep)),size=4)+
  facet_wrap( ~ TimePoint, ncol = 3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title ="Pa->Sa")+
  theme_cowplot()

ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Sa->Sa"),aes(x=Distance,y=MeanPropOcc, color = Strain, shape =factor(BioRep) ),size=4)+
  facet_wrap( ~ TimePoint, ncol = 3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title ="Sa->Sa")+
  theme_cowplot()
```

```{r Plot comparing prop oc between time points}
ggplot()+ 
  geom_point(data = combinedData %>% filter(Relation == "Pa->Pa"), aes(x=Distance,y=MeanPropOcc, color = TimePoint) ,size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap(BioRep ~ Strain, ncol = 2)+
  labs(title ="Pa->Pa")+
  theme_cowplot()
ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Sa->Pa"), aes(x=Distance,y=MeanPropOcc, color = TimePoint) ,size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap(BioRep ~ Strain, ncol = 2)+
  labs(title ="Sa->Pa")+
  theme_cowplot()
ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Pa->Sa"), aes(x=Distance,y=MeanPropOcc, color = TimePoint) ,size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap(BioRep ~ Strain, ncol = 2)+
  labs(title ="Pa->Sa")+
  theme_cowplot() 
ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Sa->Sa"), aes(x=Distance,y=MeanPropOcc, color = TimePoint) ,size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap(BioRep ~ Strain, ncol = 2)+
  labs(title ="Sa->Sa")+
  theme_cowplot()
```

```{r Calculating empty space}
combinedDataEmptySpaceStaph<-combinedData %>% filter(source %in% c("finalRG", "finalRR")) %>% group_by(Distance, TimePoint, Strain, BioRep, RunRep) %>% mutate(EmptySpace =1-MeanPropOcc[source == "finalRG"]-MeanPropOcc[source == "finalRR"])

combinedDataEmptySpacePa<-combinedData %>% filter(source %in% c("finalGG", "finalGR")) %>% group_by(Distance, TimePoint, Strain, BioRep, RunRep) %>% mutate(EmptySpace =1-MeanPropOcc[source == "finalGG"]-MeanPropOcc[source == "finalGR"])

combinedDataEmptySpace<-bind_rows(combinedDataEmptySpaceStaph,combinedDataEmptySpacePa)
```

```{r Plot Empty space per condition}
ggplot()+
  geom_point(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Pa"),aes(x=Distance,y=EmptySpace, color = Strain, shape =factor(BioRep)),size=4)+
  labs(title ="Empty Space Pa")+
  facet_wrap(~ TimePoint, ncol = 3)+
  theme_cowplot()

ggplot()+
  geom_point(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Sa"),aes(x=Distance,y=EmptySpace, color = Strain, shape =factor(BioRep)),size=4)+
  labs(title ="Empty Space Sa")+
  facet_wrap(~ TimePoint, ncol = 3)+
  theme_cowplot()

```
```{r Plot Empty space over time}
ggplot()+
  geom_point(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Pa"),aes(x=Distance,y=EmptySpace, color = TimePoint),size=4)+
  labs(title ="Empty Space Pa")+
  facet_wrap(BioRep~ Strain , ncol = 2)+
  theme_cowplot()

ggplot()+
  geom_point(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Sa"),aes(x=Distance,y=EmptySpace, color = TimePoint),size=4)+
  labs(title ="Empty Space Sa")+
  facet_wrap(BioRep~ Strain, ncol = 2)+
  theme_cowplot()

```
```{r Plot occupancy based on focus}
ggplot()+
  geom_line(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Pa" & Strain == "wt"),aes(x=Distance,y=MeanPropOcc, color = Interaction),size=1)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title ="Occupied Space around Pa wt",subtitle = "Over time, per biological replicate")+
  facet_wrap(BioRep ~ Time , ncol = 6)+
  theme_minimal_grid()

ggplot()+
  geom_line(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Pa" & Strain == "mut"),aes(x=Distance,y=MeanPropOcc, color = Interaction),size=1)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title ="Occupied Space around Pa mut",subtitle = "Over time, per biological replicate")+
  facet_wrap(BioRep ~ Time , ncol = 6)+
  theme_minimal_grid()

ggplot()+
  geom_line(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Sa" & Strain == "wt"),aes(x=Distance,y=MeanPropOcc, color = Interaction),size=1)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title ="Occupied Space around Sa wt",subtitle = "Over time, per biological replicate")+
  facet_wrap(BioRep ~ Time , ncol = 6)+
  theme_minimal_grid()

ggplot()+
  geom_line(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Sa" & Strain == "mut"),aes(x=Distance,y=MeanPropOcc, color = Interaction),size=1)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title ="Occupied Space around Sa mut",subtitle = "Over time, per biological replicate")+
  facet_wrap(BioRep ~ Time , ncol = 6)+
  theme_minimal_grid()

```

```{r Plot empty space and proportional occupancy*}
#This is a good idea, but I will get back to it later @3/25/20
ggplot()+
  geom_point(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Sa"),aes(x=Distance,y=MeanPropOcc, color = Strain, shape = Interaction),size=4)+
  geom_line(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Sa"),aes(x=Distance,y=EmptySpace, color = Strain),size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap(~ TimePoint, ncol = 3)+
  labs(title ="title")+
  theme_cowplot()

```
# Statistical exploration of the data.
Three individual runs were performed independently on each of the experiments. The output of each of them is the mean proportional occupancy and it's SD. So, if the runs follow a normal distribution, their variance should be similar.
```{r Plot PropOc mean and SD}
PlotPropOcMeanSD <- function(df, name) {
  plot1<- ggplot()+
    geom_linerange(data=df%>% filter(RunRep==1), aes(x=Distance,ymax=MeanPropOcc+SDPropOcc,ymin=MeanPropOcc-SDPropOcc, color = Relation),size=1)+
    facet_wrap(Time~Relation, ncol = 4)+
    scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title =paste0(name, " proportional occupancy mean +/- SD"))+
    theme_minimal_grid()+
    background_grid()
  ggsave(paste0(name,".png"),plot = plot1, width = 17, height = 10)
  }

PlotPropOcMeanSD(wt1, "wt1")
PlotPropOcMeanSD(wt2, "wt2")
PlotPropOcMeanSD(wt3, "wt3")
PlotPropOcMeanSD(mut1, "mut1")
PlotPropOcMeanSD(mut2, "mut2")
PlotPropOcMeanSD(mut3, "mut3")

PlotPropOcAllMeanSD <- function(relation) {
  plot1<- ggplot()+
    geom_linerange(data=combinedData %>% filter(Relation ==relation), 
                   aes(x=Distance,ymax=MeanPropOcc+SDPropOcc,ymin=MeanPropOcc-SDPropOcc, color = Strain, linetype =factor(BioRep), group = Experiment),
                   position = position_dodge(width =.8), size=.7)+
    facet_wrap(~Time, ncol = 3)+
    scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title =paste0(relation, " proportional occupancy mean +/- SD"), linetype = "Biological Replicate")+
    theme_minimal_grid()+
    background_grid()+
    theme(plot.title = element_text(hjust = 0.5))
  ggsave(paste0(relation,".png"),plot = plot1, width = 20, height = 10)
}
PlotPropOcAllMeanSD("Pa->Pa")
PlotPropOcAllMeanSD("Pa->Sa")
PlotPropOcAllMeanSD("Sa->Sa")
PlotPropOcAllMeanSD("Sa->Pa")
```
Another metric used is the coefficient of variation, CV, which is the ratio of the standard deviation to the mean.So, let's also visualize that.
```{r Plot CV} 
PlotCVrelation <- function(relation){
  plot1<-ggplot()+ 
    geom_point(data=combinedData %>% filter(Relation =="Pa->Pa"), 
               aes(x=Distance,y = CV, color = Strain, shape = factor(BioRep), group = Experiment),
               position = position_dodge(width =.8), size=4)+
    geom_hline(yintercept = 1, linetype = "dashed")+
    facet_wrap( ~ Time, ncol = 3)+
    scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title =paste0(relation, " proportional occupancy Coefficient of variation (SD/mean)"), linetype = "Biological Replicate")+
    theme_minimal_grid()+
    background_grid()+ 
    theme(plot.title = element_text(hjust = 0.5)) 
  ggsave(paste0(relation,"-CV.eps"),plot = plot1, width = 20, height = 10)
  ggsave(paste0(relation,"-CV.png"),plot = plot1, width = 20, height = 10)
  }
PlotCVrelation("Pa->Pa")
PlotCVrelation("Pa->Sa")
PlotCVrelation("Sa->Sa")
PlotCVrelation("Sa->Pa")
```
# Importing whole data sets to identify distribution of data
```{r read whole data sets}
wt1.1Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/wt1-1000-1.csv"),header = T)
wt2.1Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/wt2-1000-1.csv"),header = T)
wt3.1Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/wt3-1000-1.csv"),header = T)

mut1.1Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/mut1-1000-1.csv"),header = T)
mut2.1Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/mut2-1000-1.csv"),header = T)
mut3.1Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/mut3-1000-1.csv"),header = T)
```

```{r functions to get propOc frequency tables}
HistogramDistance <- function(df) {
  histogram<- hist(df$PropOc, breaks = seq.int(0,1.02,.01)) 
  #histogram<- hist(df$PropOc) 
  output<-data.frame(histogram$mids,histogram$counts)
  colnames(output)<-c("Value", "Frequency")
  output<-output %>% mutate(Distance = df$Distance[[1]]) %>% filter(!Frequency==0)
  return(output)
}
GetHistPropOc <- function(df,timePoint, sources) {
  data<-df %>%  filter(TimePoint==timePoint) %>% filter(source==sources) %>% group_by(Distance) %>% filter(PropOc <1.02)
  dataList<-group_split(data)
  histogramList<-map(dataList, HistogramDistance)
  output<-bind_rows(histogramList)
  output<-output %>% mutate(Source = sources)
  return(output)
}
PlotAllPropOcData <- function(df, relation, timepoint) {
  ggplot()+
    geom_point(data = df,aes(x=Value,y=Frequency, color = Distance),size=2)+
    scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    #scale_x_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title =paste( relation, timepoint))+
    #scale_color_viridis(option = "D")+
    scale_color_gradientn(colours = rainbow(5))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
}
GetAllHistPropOc <- function(df, timepoint){
  data.GG<-GetHistPropOc(df, timepoint, "finalGG")
  data.RG<-GetHistPropOc(df, timepoint, "finalRG")
  data.RR<-GetHistPropOc(df, timepoint, "finalRR")
  data.GR<-GetHistPropOc(df, timepoint, "finalGR")
return(list(data.GG,data.RR,data.RG,data.GR))
}
testGetAllHistPropOc<-GetAllHistPropOc(mut3.1Whole,5)

wt1.GG<-GetHistPropOc(wt1.1Whole, 0, "finalGG")
wt1.RG<-GetHistPropOc(wt1.1Whole, 0, "finalRG")
wt1.RR<-GetHistPropOc(wt1.1Whole, 0, "finalRR")
wt1.GR<-GetHistPropOc(wt1.1Whole, 0, "finalGR")

wt1.GG.5<-GetHistPropOc(wt1.1Whole, 5, "finalGG")
wt1.RG.5<-GetHistPropOc(wt1.1Whole, 5, "finalRG")
wt1.RR.5<-GetHistPropOc(wt1.1Whole, 5, "finalRR")
wt1.GR.5<-GetHistPropOc(wt1.1Whole, 5, "finalGR")

PlotAllPropOcData(wt1.GG)
PlotAllPropOcData(wt1.GR)
PlotAllPropOcData(wt1.RR)
PlotAllPropOcData(wt1.RG)

PlotAllPropOcData(wt1.GG.5,5, "PA->PA")
PlotAllPropOcData(wt1.GR.5,5, "PA->SA")
PlotAllPropOcData(wt1.RR.5,5, "SA->SA")
PlotAllPropOcData(wt1.RG.5,5, "SA->PA")

```