### 2020-04-28 R Analysis Scripts/PropOc Data 3D.Rmd
Up to date with Enrichment data from 50 percentile density from whole distributions

---
title: "PropOcData2"
author: "Juan P Barraza"
date: "3/5/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(pixmap)
library(rtiff)
library(reshape2)
library(gplots)
library(gdata)
library(plyr)
library(dplyr)
library(ggplot2)
library(cowplot)
library(tidyr)
library(purrr)
library(plotrix)
library(ggsignif)
library(rdist)
library("vcd")#this is fro fitting models to data
library(viridis)#color palette
library(gganimate)#gifs for ggplot

lseq <- function(from, to, length.out) {
  # logarithmic spaced sequence
  # blatantly stolen from library("emdbook"), because need only this
  exp(seq(log(from), log(to), length.out = length.out))}

source('/Users/jupabago/Documents/Whiteley/PROJECTS/Rpackage/PropOc Package/R/PropOcFunctions2.0.R') 
#using the new 3D script with capped spheres.
```

```{r functions to get PropOc Data}
GetPropOcData<-function(sample, timepoints, positions){
  folder<- paste0('/Volumes/Seagate Backup Plus Drive/Good images/',sample,'/imagesSubtractRed')
  growthDf<-data.frame()#data frame to add all timepoints and positions
  fileList<-list.files(path = folder, full.names = TRUE)
  print(folder)
  for (position in 0:positions){
    currentPosition<-fileList[grep(paste0("p",position,"_"),fileList)]
    print(paste0('analyzing position ', position))
    for (timepoint in 0:timepoints){
      currentTimePoint<-currentPosition[grep(paste0("t",GetSlice(timepoint),"_"),currentPosition)]
      dataTimePoint<-NormRawPipeline(SlicesTo3D(currentTimePoint,"red"),SlicesTo3D(currentTimePoint,"green"), 
                                     xydimSearch=114,zdimSearch = 70, sampleSize=1000,xyRealDim=0.264,
                                     zRealDim = 0.44, pipeBinSize = 1)
      dataTimePoint$TimePoint <- timepoint
      dataTimePoint$Position <- position
      growthDf<-rbind(growthDf,dataTimePoint)
      }
    }
  return(growthDf)
  }

GetSlice<-function(idx){
if (idx<10){num = paste0(0,idx)}
  else
    {num = idx}
  return(num)}
```
```{r run functions}
wt1<-GetPropOcData("3-13-19",6,2)
wt2<-GetPropOcData("3-19-19",6,2)
wt3<-GetPropOcData("4-17-19",6,2)
mut1<-GetPropOcData("4-24-19",6,2)
mut2<-GetPropOcData("4-25-19",6,2)
mut3<-GetPropOcData("5-8-19",6,2)

```
```{r Read data files}

Get3DPropOcData<-function(strain, biologicalRep, experiment){
  propOcData1<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataRedSub1/", strain,biologicalRep,"-1000-",1,".csv"),header = T)
  propOcData1<-propOcData1 %>% mutate(Strain = strain,BioRep = biologicalRep, RunRep = 1)
  propOcData2<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataRedSub1/", strain,biologicalRep,"-1000-",2,".csv"),header = T)
  propOcData2<-propOcData2 %>% mutate(Strain = strain,BioRep = biologicalRep, RunRep = 2)
  propOcData3<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataRedSub1/", strain,biologicalRep,"-1000-",3,".csv"),header = T)
  propOcData3<-propOcData3 %>% mutate(Strain = strain,BioRep = biologicalRep, RunRep = 3)
  allData<-bind_rows(propOcData1,propOcData2,propOcData3)
  allData<-allData %>%  mutate(CV = SDPropOcc/MeanPropOcc) %>% mutate(Experiment = experiment)
  allData<-allData %>% 
    mutate(Relation = recode(source, "finalGG" = "Pa->Pa","finalRG" = "Sa->Pa","finalRR" = "Sa->Sa","finalGR" = "Pa->Sa")) %>%
    mutate(Condition = recode(Position, "2" = "Pa:Sa 1:1")) %>%
    mutate(Interaction = recode(source, "finalGG" = "Self","finalRG" = "Other","finalRR" = "Self","finalGR" = "Other")) %>%
    mutate(FocusSpecies = recode(source, "finalGG" = "Pa","finalRG" = "Sa","finalRR" = "Sa","finalGR" = "Pa")) %>%
    mutate(TargetSpecies = recode(source, "finalGG" = "Pa","finalRG" = "Pa","finalRR" = "Sa","finalGR" = "Sa")) %>% 
    mutate(Time = recode(TimePoint, "0" = "Hour 1","1" = "Hour 2","2" = "Hour 3","3" = "Hour 4","4" = "Hour 5","5" = "Hour 6"))

  allData<-allData %>% filter(TimePoint<6 & Distance<30)#last distance is always off for some reason =(
  }

wt1<-Get3DPropOcData("wt", 1, "wt1")
wt2<-Get3DPropOcData("wt", 2, "wt2")
wt3<-Get3DPropOcData("wt", 3, "wt3")
wt3<-wt3 %>% filter(!RunRep==2|!TimePoint==1)

mut1<-Get3DPropOcData("mut", 1, "mut1")
mut2<-Get3DPropOcData("mut", 2, "mut2")
mut3<-Get3DPropOcData("mut", 3, "mut3")

combinedData<-bind_rows(wt1,wt2,wt3,mut1,mut2,mut3)
combinedData<-combinedData %>% filter(RunRep==1)#I just saw that theyr are all the same, so no point on dragging that much data
```
```{r checking the replicate runs}
PlotPropOcAll <- function(df, name) {
  ggplot()+
    geom_point(data=df, aes(x=Distance,y=MeanPropOcc, color = Relation, shape = factor(RunRep)),size=4)+
    geom_errorbar(data=df, aes(x=Distance,ymin=MeanPropOcc-SDPropOcc,ymax=MeanPropOcc+SDPropOcc, color = Relation, linetype = factor(RunRep)),size=4)+
    facet_wrap(TimePoint~Relation, ncol = 4)+
    scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title =name)+
    theme_cowplot()+
    background_grid()
  }
PlotPropOcAll(wt1,"wt1")
PlotPropOcAll(wt2,"wt2")
PlotPropOcAll(wt3,"wt3")
PlotPropOcAll(mut1,"mut1 ")
PlotPropOcAll(mut2,"mut2")
PlotPropOcAll(mut3,"mut3")
```

```{r Plot comparison of prop oc between Strains}
ggplot()+ 
  geom_point(data = combinedData %>% filter(Relation == "Pa->Pa"),aes(x=Distance,y=MeanPropOcc, color = Strain, shape =factor(BioRep) ),size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap( ~ TimePoint, ncol = 3)+
  labs(title ="Pa->Pa")+  
  theme_cowplot()

ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Sa->Pa"),aes(x=Distance,y=MeanPropOcc, color = Strain, shape =factor(BioRep) ),size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap( ~ TimePoint, ncol = 3)+
  labs(title ="Sa->Pa")+
  theme_cowplot()

ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Pa->Sa"),aes(x=Distance,y=MeanPropOcc, color = Strain, shape =factor(BioRep)),size=4)+
  facet_wrap( ~ TimePoint, ncol = 3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title ="Pa->Sa")+
  theme_cowplot()

ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Sa->Sa"),aes(x=Distance,y=MeanPropOcc, color = Strain, shape =factor(BioRep) ),size=4)+
  facet_wrap( ~ TimePoint, ncol = 3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title ="Sa->Sa")+
  theme_cowplot()
```

```{r Plot comparing prop oc between time points}
ggplot()+ 
  geom_point(data = combinedData %>% filter(Relation == "Pa->Pa"), aes(x=Distance,y=MeanPropOcc, color = TimePoint) ,size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap(BioRep ~ Strain, ncol = 2)+
  labs(title ="Pa->Pa")+
  theme_cowplot()
ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Sa->Pa"), aes(x=Distance,y=MeanPropOcc, color = TimePoint) ,size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap(BioRep ~ Strain, ncol = 2)+
  labs(title ="Sa->Pa")+
  theme_cowplot()
ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Pa->Sa"), aes(x=Distance,y=MeanPropOcc, color = TimePoint) ,size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap(BioRep ~ Strain, ncol = 2)+
  labs(title ="Pa->Sa")+
  theme_cowplot() 
ggplot()+
  geom_point(data = combinedData %>% filter(Relation == "Sa->Sa"), aes(x=Distance,y=MeanPropOcc, color = TimePoint) ,size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap(BioRep ~ Strain, ncol = 2)+
  labs(title ="Sa->Sa")+
  theme_cowplot()
```

```{r Calculating empty space}
combinedDataEmptySpaceStaph<-combinedData %>% filter(source %in% c("finalRG", "finalRR")) %>% group_by(Distance, TimePoint, Strain, BioRep, RunRep) %>% mutate(EmptySpace =1-MeanPropOcc[source == "finalRG"]-MeanPropOcc[source == "finalRR"])

combinedDataEmptySpacePa<-combinedData %>% filter(source %in% c("finalGG", "finalGR")) %>% group_by(Distance, TimePoint, Strain, BioRep, RunRep) %>% mutate(EmptySpace =1-MeanPropOcc[source == "finalGG"]-MeanPropOcc[source == "finalGR"])

combinedDataEmptySpace<-bind_rows(combinedDataEmptySpaceStaph,combinedDataEmptySpacePa)
```

```{r Plot Empty space per condition}
ggplot()+
  geom_point(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Pa"),aes(x=Distance,y=EmptySpace, color = Strain, shape =factor(BioRep)),size=4)+
  labs(title ="Empty Space Pa")+
  facet_wrap(~ TimePoint, ncol = 3)+
  theme_cowplot()

ggplot()+
  geom_point(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Sa"),aes(x=Distance,y=EmptySpace, color = Strain, shape =factor(BioRep)),size=4)+
  labs(title ="Empty Space Sa")+
  facet_wrap(~ TimePoint, ncol = 3)+
  theme_cowplot()

```
```{r Plot Empty space over time}
ggplot()+
  geom_point(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Pa"),aes(x=Distance,y=EmptySpace, color = TimePoint),size=4)+
  labs(title ="Empty Space Pa")+
  facet_wrap(BioRep~ Strain , ncol = 2)+
  theme_cowplot()

ggplot()+
  geom_point(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Sa"),aes(x=Distance,y=EmptySpace, color = TimePoint),size=4)+
  labs(title ="Empty Space Sa")+
  facet_wrap(BioRep~ Strain, ncol = 2)+
  theme_cowplot()

```
```{r Plot occupancy based on focus}
ggplot()+
  geom_line(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Pa" & Strain == "wt"),aes(x=Distance,y=MeanPropOcc, color = Interaction),size=1)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title ="Occupied Space around Pa wt",subtitle = "Over time, per biological replicate")+
  facet_wrap(BioRep ~ Time , ncol = 6)+
  theme_minimal_grid()

ggplot()+
  geom_line(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Pa" & Strain == "mut"),aes(x=Distance,y=MeanPropOcc, color = Interaction),size=1)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title ="Occupied Space around Pa mut",subtitle = "Over time, per biological replicate")+
  facet_wrap(BioRep ~ Time , ncol = 6)+
  theme_minimal_grid()

ggplot()+
  geom_line(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Sa" & Strain == "wt"),aes(x=Distance,y=MeanPropOcc, color = Interaction),size=1)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title ="Occupied Space around Sa wt",subtitle = "Over time, per biological replicate")+
  facet_wrap(BioRep ~ Time , ncol = 6)+
  theme_minimal_grid()

ggplot()+
  geom_line(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Sa" & Strain == "mut"),aes(x=Distance,y=MeanPropOcc, color = Interaction),size=1)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title ="Occupied Space around Sa mut",subtitle = "Over time, per biological replicate")+
  facet_wrap(BioRep ~ Time , ncol = 6)+
  theme_minimal_grid()

```

```{r Plot empty space and proportional occupancy*}
#This is a good idea, but I will get back to it later @3/25/20
ggplot()+
  geom_point(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Sa"),aes(x=Distance,y=MeanPropOcc, color = Strain, shape = Interaction),size=4)+
  geom_line(data = combinedDataEmptySpace %>% filter(FocusSpecies=="Sa"),aes(x=Distance,y=EmptySpace, color = Strain),size=4)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap(~ TimePoint, ncol = 3)+
  labs(title ="title")+
  theme_cowplot()

```
# Statistical exploration of the data.
Three individual runs were performed independently on each of the experiments. The output of each of them is the mean proportional occupancy and it's SD. So, if the runs follow a normal distribution, their variance should be similar.
```{r Plot PropOc mean and SD}
PlotPropOcMeanSD <- function(df, name) {
  plot1<- ggplot()+
    geom_linerange(data=df%>% filter(RunRep==1), aes(x=Distance,ymax=MeanPropOcc+SDPropOcc,ymin=MeanPropOcc-SDPropOcc, color = Relation),size=1)+
    facet_wrap(Time~Relation, ncol = 4)+
    scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title =paste0(name, " proportional occupancy mean +/- SD"))+
    theme_minimal_grid()+
    background_grid()
  ggsave(paste0(name,".png"),plot = plot1, width = 17, height = 10)
  }

PlotPropOcMeanSD(wt1, "wt1")
PlotPropOcMeanSD(wt2, "wt2")
PlotPropOcMeanSD(wt3, "wt3")
PlotPropOcMeanSD(mut1, "mut1")
PlotPropOcMeanSD(mut2, "mut2")
PlotPropOcMeanSD(mut3, "mut3")

PlotPropOcAllMeanSD <- function(relation) {
  plot1<- ggplot()+
    geom_linerange(data=combinedData %>% filter(Relation ==relation), 
                   aes(x=Distance,ymax=MeanPropOcc+SDPropOcc,ymin=MeanPropOcc-SDPropOcc, color = Strain, linetype =factor(BioRep), group = Experiment),
                   position = position_dodge(width =.8), size=.7)+
    facet_wrap(~Time, ncol = 3)+
    scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title =paste0(relation, " proportional occupancy mean +/- SD"), linetype = "Biological Replicate")+
    theme_minimal_grid()+
    background_grid()+
    theme(plot.title = element_text(hjust = 0.5))
  ggsave(paste0(relation,".png"),plot = plot1, width = 20, height = 10)
}
PlotPropOcAllMeanSD("Pa->Pa")
PlotPropOcAllMeanSD("Pa->Sa")
PlotPropOcAllMeanSD("Sa->Sa")
PlotPropOcAllMeanSD("Sa->Pa")
```
Another metric used is the coefficient of variation, CV, which is the ratio of the standard deviation to the mean.So, let's also visualize that.
```{r Plot CV} 
PlotCVrelation <- function(relation){
  plot1<-ggplot()+ 
    geom_point(data=combinedData %>% filter(Relation =="Pa->Pa"), 
               aes(x=Distance,y = CV, color = Strain, shape = factor(BioRep), group = Experiment),
               position = position_dodge(width =.8), size=4)+
    geom_hline(yintercept = 1, linetype = "dashed")+
    facet_wrap( ~ Time, ncol = 3)+
    scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title =paste0(relation, " proportional occupancy Coefficient of variation (SD/mean)"), linetype = "Biological Replicate")+
    theme_minimal_grid()+
    background_grid()+ 
    theme(plot.title = element_text(hjust = 0.5)) 
  ggsave(paste0(relation,"-CV.eps"),plot = plot1, width = 20, height = 10)
  ggsave(paste0(relation,"-CV.png"),plot = plot1, width = 20, height = 10)
  }
PlotCVrelation("Pa->Pa")
PlotCVrelation("Pa->Sa")
PlotCVrelation("Sa->Sa")
PlotCVrelation("Sa->Pa")
```
# Importing whole data sets to identify distribution of data
```{r read whole data sets}
wt1.1Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/wt1-1000-1.csv"),header = T)
wt2.1Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/wt2-1000-1.csv"),header = T)
wt3.1Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/wt3-1000-1.csv"),header = T)

mut1.1Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/mut1-1000-1.csv"),header = T)
mut2.1Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/mut2-1000-1.csv"),header = T)
mut3.1Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/mut3-1000-1.csv"),header = T)

wt1.2Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/wt1-1000-2.csv"),header = T)
wt2.2Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/wt2-1000-2.csv"),header = T)
wt3.2Whole<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/PropOcDataAll/wt3-1000-2.csv"),header = T)

```

```{r Whole PropOc frequency functions }
GetAllHistPropOc <- function(df, timepoint){
  data.GG<-GetHistPropOc(df, timepoint, "finalGG")
  data.RG<-GetHistPropOc(df, timepoint, "finalRG")
  data.RR<-GetHistPropOc(df, timepoint, "finalRR")
  data.GR<-GetHistPropOc(df, timepoint, "finalGR")
  allData<- bind_rows(data.GG,data.RR,data.RG,data.GR)
return(allData)
}

GetHistPropOc <- function(df,timePoint, sources) {
  data<-df %>%  filter(TimePoint==timePoint) %>% filter(source==sources) %>% group_by(Distance) %>% filter(PropOc <1.02)
  dataList<-group_split(data)
  histogramList<-map(dataList, HistogramDistance)
  output<-bind_rows(histogramList)
  output<-output %>% mutate(Source = sources)
  return(output)
}
HistogramDistance <- function(df) {
  #histogram<- hist(df$PropOc, breaks = seq.int(0,1.02,.01)) 
  #histogram<- hist(df$PropOc) 
  #histogram<- hist(df$PropOc, breaks = c(0,lseq(.001,1.03,100)))
  histogram<- hist(df$PropOc, breaks = c(0,lseq(.0001,1.03,100)))
  output<-data.frame(histogram$mids,histogram$counts, histogram$density)
  colnames(output)<-c("Value", "Frequency", "Density")
  output<-output %>% mutate(Distance = df$Distance[[1]]) %>% filter(!Frequency==0)
  return(output)
}

#Get all of the distributions
#below are for first run:
wt1.1Whole.5<-GetAllHistPropOc(wt1.1Whole,5)
wt2.1Whole.5<-GetAllHistPropOc(wt2.1Whole,5)
wt3.1Whole.5<-GetAllHistPropOc(wt3.1Whole,5)
mut1.1Whole.5<-GetAllHistPropOc(mut1.1Whole,5)
mut2.1Whole.5<-GetAllHistPropOc(mut2.1Whole,5)
mut3.1Whole.5<-GetAllHistPropOc(mut3.1Whole,5)
#second run
wt1.2Whole.5<-GetAllHistPropOc(wt1.2Whole,5)
wt2.2Whole.5<-GetAllHistPropOc(wt2.2Whole,5)
wt3.2Whole.5<-GetAllHistPropOc(wt3.2Whole,5)

#log sequence
wt1.1Whole.5log<-GetAllHistPropOc(wt1.1Whole,5)  
wt2.1Whole.5log<-GetAllHistPropOc(wt2.1Whole,5)
wt3.1Whole.5log<-GetAllHistPropOc(wt3.1Whole,5)
mut1.1Whole.5log<-GetAllHistPropOc(mut1.1Whole,5)
mut2.1Whole.5log<-GetAllHistPropOc(mut2.1Whole,5)
mut3.1Whole.5log<-GetAllHistPropOc(mut3.1Whole,5)

#log starting at .0001
wt1.1Whole.5log2<-GetAllHistPropOc(wt1.1Whole,5)  
wt2.1Whole.5log2<-GetAllHistPropOc(wt2.1Whole,5)
wt3.1Whole.5log2<-GetAllHistPropOc(wt3.1Whole,5)
mut1.1Whole.5log2<-GetAllHistPropOc(mut1.1Whole,5)
mut2.1Whole.5log2<-GetAllHistPropOc(mut2.1Whole,5)
mut3.1Whole.5log2<-GetAllHistPropOc(mut3.1Whole,5)
```

```{r Whole PropOc frequency Figures}
PlotAllPropOcData <- function(df, timepoint, relation ) {
  plot<-ggplot()+ 
    geom_point(data = df,aes(x=Value,y=Frequency, color = Distance),size=2)+
    scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    #scale_x_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title =paste( relation, timepoint))+
    #scale_color_viridis(option = "D")+
    scale_color_gradientn(colours = rainbow(5))+
    xlim(0,1)+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
  name<-paste0(relation, ".hour", timepoint,".png")
  ggsave(paste0(name),plot = plot, width = 17, height = 10)
}

PlotAllPropOcData(wt1.1Whole.5 %>% filter (Source == "finalGG"), 5, "Wt 1-1 PA->PA")
PlotAllPropOcData(wt2.1Whole.5 %>% filter (Source == "finalGG"), 5, "Wt 2-1 PA->PA")
PlotAllPropOcData(wt3.1Whole.5 %>% filter (Source == "finalGG"), 5, "Wt 3-1 PA->PA")

PlotAllPropOcData(wt1.2Whole.5 %>% filter (Source == "finalGG"), 5, "Wt 1-2 PA->PA")
PlotAllPropOcData(wt2.2Whole.5 %>% filter (Source == "finalGG"), 5, "Wt 2-2 PA->PA")
PlotAllPropOcData(wt3.2Whole.5 %>% filter (Source == "finalGG"), 5, "Wt 3-2 PA->PA")

PlotAllPropOcData(mut1.1Whole.5 %>% filter (Source == "finalGG"), 5, "Mut 1-1 PA->PA")
PlotAllPropOcData(mut2.1Whole.5 %>% filter (Source == "finalGG"), 5, "Mut 2-1 PA->PA")
PlotAllPropOcData(mut3.1Whole.5 %>% filter (Source == "finalGG"), 5, "Mut 3-1 PA->PA")

PlotAllPropOcData(wt1.1Whole.5 %>% filter (Source == "finalRG"), 5, "Wt 1-1 SA->PA")
PlotAllPropOcData(wt2.1Whole.5 %>% filter (Source == "finalRG"), 5, "Wt 2-1 SA->PA")
PlotAllPropOcData(wt3.1Whole.5 %>% filter (Source == "finalRG"), 5, "Wt 3-1 SA->PA")

PlotAllPropOcData(wt1.2Whole.5 %>% filter (Source == "finalRG"), 5, "Wt 1-2 SA->PA")
PlotAllPropOcData(wt2.2Whole.5 %>% filter (Source == "finalRG"), 5, "Wt 2-2 SA->PA")
PlotAllPropOcData(wt3.2Whole.5 %>% filter (Source == "finalRG"), 5, "Wt 3-2 SA->PA")

PlotAllPropOcData(mut1.1Whole.5 %>% filter (Source == "finalRG"), 5, "Mut1-1 SA->PA")
PlotAllPropOcData(mut2.1Whole.5 %>% filter (Source == "finalRG"), 5, "Mut2-1 SA->PA")
PlotAllPropOcData(mut3.1Whole.5 %>% filter (Source == "finalRG"), 5, "Mut3-1 SA->PA")

PlotAllPropOcData(wt1.1Whole.5 %>% filter (Source == "finalRR"), 5, "Wt 1-1 SA->SA")
PlotAllPropOcData(wt2.1Whole.5 %>% filter (Source == "finalRR"), 5, "Wt 2-1 SA->SA")
PlotAllPropOcData(wt3.1Whole.5 %>% filter (Source == "finalRR"), 5, "Wt 3-1 SA->SA")

PlotAllPropOcData(mut1.1Whole.5 %>% filter (Source == "finalRR"), 5, "Mut1-1 SA->SA")
PlotAllPropOcData(mut2.1Whole.5 %>% filter (Source == "finalRR"), 5, "Mut2-1 SA->SA")
PlotAllPropOcData(mut3.1Whole.5 %>% filter (Source == "finalRR"), 5, "Mut3-1 SA->SA")

PlotAllPropOcData(wt1.1Whole.5 %>% filter (Source == "finalGR"), 5, "Wt 1-1 PA->SA")
PlotAllPropOcData(wt2.1Whole.5 %>% filter (Source == "finalGR"), 5, "Wt 2-1 PA->SA")
PlotAllPropOcData(wt3.1Whole.5 %>% filter (Source == "finalGR"), 5, "Wt 3-1 PA->SA")

PlotAllPropOcData(mut1.1Whole.5 %>% filter (Source == "finalGR"), 5, "Mut1-1 PA->SA")
PlotAllPropOcData(mut2.1Whole.5 %>% filter (Source == "finalGR"), 5, "Mut2-1 PA->SA")
PlotAllPropOcData(mut3.1Whole.5 %>% filter (Source == "finalGR"), 5, "Mut3-1 PA->SA")

#Honestly, they all look pretty much the same...
#let's try doing the bins in an exponential function...
PlotAllPropOcDatalog <- function(df, timepoint, relation ) {
  plot<-ggplot()+ 
    geom_point(data = df,aes(x=Value,y=Frequency, color = Distance),size=2)+
    scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)),limits = (c(1,1001)))+
    #scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)),limits = (c(0.0005,1)))+
    labs(title =paste( relation, timepoint))+
    #scale_color_viridis(option = "D")+
    scale_color_gradientn(colours = rainbow(5))+
    #xlim(0,1)+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
  name<-paste0(relation, ".hour", timepoint,"(log).png")
  ggsave(paste0(name),plot = plot, width = 17, height = 10)
}

PlotAllPropOcDatalog(wt1.1Whole.5log %>% filter (Source == "finalRG"), 5, "Wt 1-1 SA->PA")
PlotAllPropOcDatalog(wt2.1Whole.5log %>% filter (Source == "finalRG"), 5, "Wt 2-1 SA->PA")
PlotAllPropOcDatalog(wt3.1Whole.5log %>% filter (Source == "finalRG"), 5, "Wt 3-1 SA->PA")

PlotAllPropOcDatalog(mut1.1Whole.5log %>% filter (Source == "finalRG"), 5, "Mut1-1 SA->PA")
PlotAllPropOcDatalog(mut2.1Whole.5log %>% filter (Source == "finalRG"), 5, "Mut2-1 SA->PA")
PlotAllPropOcDatalog(mut3.1Whole.5log %>% filter (Source == "finalRG"), 5, "Mut3-1 SA->PA")

PlotAllPropOcDatalog(wt1.1Whole.5log %>% filter (Source == "finalGR"), 5, "Wt 1-1 PA->SA")
PlotAllPropOcDatalog(wt2.1Whole.5log %>% filter (Source == "finalGR"), 5, "Wt 2-1 PA->SA")
PlotAllPropOcDatalog(wt3.1Whole.5log %>% filter (Source == "finalGR"), 5, "Wt 3-1 PA->SA")

PlotAllPropOcDatalog(mut1.1Whole.5log %>% filter (Source == "finalGR"), 5, "Mut1-1 PA->SA")
PlotAllPropOcDatalog(mut2.1Whole.5log %>% filter (Source == "finalGR"), 5, "Mut2-1 PA->SA")
PlotAllPropOcDatalog(mut3.1Whole.5log %>% filter (Source == "finalGR"), 5, "Mut3-1 PA->SA")

```

```{r Whole PropOc Data wrangling}
#So, I think the first thing to do would be to get a "mode", or the highest value for frequency of proportional occupancy per distance.
#at this point I can see at least if the zone of interest is the same for every sample.

GetMaxDensity <- function(df) {#this one is essentially the same as the one below, but thos one doesnt allow repeated max density values
  df.summary<-df %>% mutate(Density = Frequency/1000) %>% #this is to make it be from 0 to 1
    group_by(Distance, Source) %>% select(Value, Distance,Density, Source) %>% summarise(DensestValue = Value [Density ==max(Density)], MaxDensity = max(Density))
  return(df.summary)
}

DensityCheck <- function(df) {
  df.summary<-df %>% 
    mutate(Density = Frequency/1000) %>% #this is to make it be from 0 to 1
    group_by(Distance, Source) %>% select(Value, Distance,Density, Source) %>% 
    mutate(MaxDensity = max(Density)) %>% 
    mutate(DensityCh = ifelse(Density ==max(Density), 1, 0 )) %>% filter(DensityCh==1) %>% ungroup()#find max density, check where it matches
  df.summary<-df.summary %>% group_by(Distance, Source) %>% mutate(DoubledValues = sum(DensityCh))#this is just to find which values were repeated
  return(df.summary)
}

ProcessDensities <- function(histograms) {
histogramNames<- c("wt1.1","wt2.1","wt3.1","mut1.1","mut2.1","mut3.1")
mappedDensities<-map(histograms,DensityCheck)
namedMappedDensities<-map2(mappedDensities, histogramNames, ~.x %>% mutate(Sample = .y))
allDensities<-bind_rows(namedMappedDensities)
allDensities<-allDensities %>% ungroup() %>% 
  mutate(Strain = recode(allDensities$Sample, wt1.1="wt",wt2.1="wt",wt3.1="wt",mut1.1="mut",mut2.1="mut",mut3.1="mut"))
return(allDensities)}

histograms<- list(wt1.1Whole.5,wt2.1Whole.5,wt3.1Whole.5,mut1.1Whole.5,mut2.1Whole.5,mut3.1Whole.5)
histogramslog<- list(wt1.1Whole.5log,wt2.1Whole.5log,wt3.1Whole.5log,mut1.1Whole.5log,mut2.1Whole.5log,mut3.1Whole.5log)
histogramslog2<- list(wt1.1Whole.5log2,wt2.1Whole.5log2,wt3.1Whole.5log2,mut1.1Whole.5log2,mut2.1Whole.5log2,mut3.1Whole.5log2)

allDensitiesFunctions<-ProcessDensities(histograms)
allDensitiesLog<-ProcessDensities(histogramslog)
allDensitiesLog2<-ProcessDensities(histogramslog2)
```

```{r Whole Prop Freq Figures}
PlotMaxDensities <- function(df) {
  ggplot()+
    geom_jitter(data = df,aes(y=Value, color=Density,x = Distance, shape = Sample),size=2)+
    labs(title ="Maximum Density")+
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    #scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    facet_wrap( ~ Source, ncol = 2)+
    scale_color_gradientn(colours = rainbow(5))+
    scale_shape_manual(values=c(0,1,2, 15,16,17))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
}
PlotMaxDensities(allDensities)
PlotMaxDensities(allDensitiesFunctions %>% filter(Source=="finalRG"))
PlotMaxDensities(allDensitiesLog)
PlotMaxDensities(allDensitiesLog2)
```
#Calculating 50 percentile density.
```{r death curves Functions and data}
getDeathCurve<-function(df){
  totalFrequency = sum(df$Frequency)
  currentFrequency = 0
  for (i in 1 : nrow(df)){
    addedFrequency<-totalFrequency-currentFrequency
    df[i,6]<-addedFrequency
    currentFrequency = currentFrequency+df[i,2]
  }
  colnames(df)<-c("Value","Frequency","Density", "Distance", "Source","Decay")
  df<-df %>% mutate (normDecay = 100*(Decay - min(Decay)) / (max(Decay) - min(Decay)))
  return(df)
  }
processData<-function(df){
  allData<-df %>% group_by(Distance, Source)
  yas<- group_split(allData)
  mappedData<-map(yas, getDeathCurve)
  boundData<-bind_rows(mappedData)
}
GetPercentiles<-function(histograms){
  histogramNames<- c("wt1.1","wt2.1","wt3.1","mut1.1","mut2.1","mut3.1")
  mappedDensities<-map(histograms,processData)
  namedMappedDensities<-map2(mappedDensities, histogramNames, ~.x %>% mutate(Sample = .y))
  allDensities<-bind_rows(namedMappedDensities)
  allDensities<-allDensities %>% ungroup() %>% 
    mutate(Strain = recode(allDensities$Sample, wt1.1="wt",wt2.1="wt",wt3.1="wt",mut1.1="mut",mut2.1="mut",mut3.1="mut"))
return(allDensities)}

PlotPercentiles <- function(df) {
  ggplot()+
    geom_point(data = df,aes(y=Value, color=Percentile50,x = Distance, shape = Sample),size=2)+
    #geom_pointrange(data = df,aes(x = Distance,y=meanValue,ymax=meanValue+SdValue,ymin=meanValue-SdValue, color=Strain),size=2)+
    labs(title ="50 percentile Density")+
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    #scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    facet_wrap( ~ Source, ncol = 2)+
    scale_color_gradientn(colours = rainbow(5))+
    scale_shape_manual(values=c(0,1,2, 15,16,17))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
  }

allDeathCurvesLog2<-GetPercentiles(histogramslog2)
#there are, I believe, at least 3 ways to find the 50 percentile. one of them is finidng the closest value, regardless of where it comes from, whether it is higher or lower than 50 percentile. The other two are either taking the closest to 50, either exclusively larger or exclusively smaller than that. That's why there are 3 ways of doing this here:

#this is the closest to absolute value 50
allDeathCurvesLog2midAbs <- allDeathCurvesLog2 %>% group_by(Sample, Distance, Source) %>% 
  mutate(Percentile50 = min(abs(normDecay - 50)))%>% 
  mutate(DensityCh = ifelse(Percentile50 ==abs(normDecay - 50), 1, 0 )) %>% filter(DensityCh==1) %>% mutate(DoubledValues = sum(DensityCh)) %>% ungroup()#find max density, check where it matches

#this is coming from the highest value
allDeathCurvesLog2topAbs <- allDeathCurvesLog2 %>% group_by(Sample, Distance, Source) %>% 
  mutate(Percentile50 = (normDecay - 50)) %>% filter (Percentile50>0) %>% filter(Percentile50==min(Percentile50)) %>%  
  mutate(DensityCh = ifelse(Percentile50 ==(normDecay - 50), 1, 0 )) %>% filter(DensityCh==1) %>% 
  mutate(DoubledValues = sum(DensityCh)) %>% ungroup()#find max density, check where it matches

#this is coming from the lowest value
allDeathCurvesLog2bottomAbs <- allDeathCurvesLog2 %>% group_by(Sample, Distance, Source) %>% 
  mutate(Percentile50 = (50 -normDecay)) %>% filter (Percentile50>0) %>% filter(Percentile50==min(Percentile50)) %>%  
  mutate(DensityCh = ifelse(Percentile50 ==(50 - normDecay), 1, 0 )) %>% filter(DensityCh==1) %>% 
  mutate(DoubledValues = sum(DensityCh)) %>% ungroup()#find max density, check where it matches

#Plotting and comparing all 3
PlotPercentiles(allDeathCurvesLog2midAbs)
PlotPercentiles(allDeathCurvesLog2topAbs)
PlotPercentiles(allDeathCurvesLog2bottomAbs)
#they all look fine, so I'll just stick with the first one

#this looks kind of ok... what is the mean and sd?
allDeathCurvesLog2midAbsStats<-allDeathCurvesLog2midAbs %>% group_by(Strain, Distance, Source) %>% mutate(LogValue = log10(Value)) %>% mutate(meanValue = mean(Value), sdValue = sd(Value), meanLogValue = mean(LogValue),sdLogValue = sd(LogValue))
PlotPercentiles(allDeathCurvesLog2midAbsStats)

ggplot()+#linear
    #geom_point(data = df,aes(y=Value, color=Percentile50,x = Distance, shape = Sample),size=2)+
    geom_pointrange(data = allDeathCurvesLog2midAbsStats,aes(x = Distance,y=meanValue,ymax=meanValue+sdValue,ymin=meanValue-sdValue, color=Strain),size=.5)+
    labs(title ="50 percentile Density")+
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    #scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    facet_wrap( ~ Source, ncol = 2)+
    #scale_color_gradientn(colours = rainbow(5))+
    #scale_shape_manual(values=c(0,1,2, 15,16,17))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))

ggplot()+#log transform
    geom_pointrange(data = allDeathCurvesLog2midAbsStats,aes(x = Distance,y=meanLogValue,ymax=meanLogValue+sdLogValue,ymin=meanLogValue-sdLogValue, color=Strain),size=.5)+
    labs(title ="50 percentile Density (Log)")+
    facet_wrap( ~ Source, ncol = 2)+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
```
#Enrichment (distance at highest proportional occupancy)
```{r Enrichment}
#find the max proportional occupancy value from the 50 percentile

allDeathCurvesLog2midAbsEnrichment<- allDeathCurvesLog2midAbs %>% group_by(Sample, Source)%>%select(Value, Distance, Source,Sample, Strain, Percentile50) %>% mutate(Enrichment = max(Value)) %>%
  mutate(EnrichmentCh = ifelse(Enrichment == Value, 1, 0 )) %>% filter(EnrichmentCh==1) %>% mutate(DoubledValues = sum(EnrichmentCh)) %>% ungroup()#find max

PlotEnrichment <- function(df) {
ggplot()+#first run
    geom_point(data = df,aes(y=Distance, x = Sample, color = Enrichment, shape = Sample),size=2)+
    labs(title ="50 percentile Enrichment")+
    #scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    #scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    facet_wrap( ~ Source, ncol = 2)+
    scale_color_gradientn(colours = rainbow(5))+
    scale_shape_manual(values=c(0,1,2, 15,16,17))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
}

PlotDataEnrichmentEmphasis <- function(wholeData, EnrichmentData) {
ggplot()+#including rest of the data
  geom_point(data = wholeData,aes(y=Value, color=Percentile50,x = Distance, shape = Sample),size=2)+
  geom_point(data = EnrichmentData,aes(y=Value, color=Percentile50,x = Distance, shape = Sample),size=5)+
  labs(title ="50 percentile Density (Enrichment emphasis)")+
  scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  #scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_wrap( ~ Source, ncol = 2)+
  scale_color_gradientn(colours = rainbow(5))+
  scale_shape_manual(values=c(0,1,2, 15,16,17))+
  theme_cowplot()+
  theme(plot.title = element_text(hjust = 0.5))
}

PlotEnrichment(allDeathCurvesLog2midAbsEnrichment)
PlotDataEnrichmentEmphasis(allDeathCurvesLog2midAbs,allDeathCurvesLog2midAbsEnrichment)

#That looks like a lot of the values are just at 0.5 distance, which I dont really care, I can make another argument about direct contact, so I'll filter those out...
allDeathCurvesLog2midAbsEnrichment2<- allDeathCurvesLog2midAbs %>% filter(!Distance ==0.5)%>% 
  group_by(Sample, Source) %>% select(Value, Distance, Source,Sample, Strain, Percentile50) %>% mutate(Enrichment = max(Value)) %>%
  mutate(EnrichmentCh = ifelse(Enrichment == Value, 1, 0 )) %>% filter(EnrichmentCh==1) %>% mutate(DoubledValues = sum(EnrichmentCh)) %>% ungroup()#find max

PlotEnrichment(allDeathCurvesLog2midAbsEnrichment2)
PlotDataEnrichmentEmphasis(allDeathCurvesLog2midAbs,allDeathCurvesLog2midAbsEnrichment2)

#This looks a little better... now get the mean per condition
allDeathCurvesLog2midAbsEnrichment2<-allDeathCurvesLog2midAbsEnrichment2 %>% group_by(Source, Sample) %>% mutate(EnrichmentDistance = mean(Distance)) %>% ungroup() %>% group_by(Source, Strain) %>% mutate(meanEnrichment = mean(EnrichmentDistance), sdEnrichment = sd(EnrichmentDistance))

ggplot()+
    geom_point(data = allDeathCurvesLog2midAbsEnrichment2,aes(y=meanEnrichment, x = Strain, color = Strain),size=4)+
    geom_errorbar(data = allDeathCurvesLog2midAbsEnrichment2,
                  aes(ymin = meanEnrichment-sdEnrichment,ymax = meanEnrichment+sdEnrichment, x = Strain, color = Strain),size=1)+
    geom_jitter(data = allDeathCurvesLog2midAbsEnrichment2,aes(y=Distance, x = Strain, color = Strain, shape = Sample),size=2)+
    labs(title ="50 percentile Enrichment by strain")+
    facet_wrap( ~ Source, ncol = 2)+
    #scale_color_gradientn(colours = rainbow(5))+
    scale_shape_manual(values=c(0,1,2, 15,16,17))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))

```