### 2020-06-24 Staph killing assays/Staph killing report.Rmd
redid figures after talking to Marvin in lab

---
title: "Staph killing assays"
author: "Juan P Barraza"
date: "1/15/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(magrittr)
library(ggplot2)
library(dplyr)
library(cowplot)
library(scales)
library(tidyr)
library(growthcurver)
library(plotrix)
```

```{r zone of inhibition data input}
diskFile<-"/Users/jupabago/Documents/Whiteley/PROJECTS/pqsL SCFM2/Staph killing assays/disk difussion assays.csv"
diskData <-read.csv(diskFile)
diskData$strain <- factor(diskData$strain, levels = c("wt", "pqsl"))
diskData <- diskData %>% mutate(val1 = value1-6,val2 = value2-6,average = (val1+val2)/2 )#this removes the diameter of the disks
diskDataStats<-diskData %>% group_by(strain, condition, bio.rep) %>% summarise(meanTech = mean(average)) %>% ungroup() %>%group_by(strain, condition) %>% summarise(meanBio = mean(meanTech), SDBio = std.error(meanTech))
diskDataStats$strain <- factor(diskDataStats$strain, levels = c("wt", "pqsl"))
diskDataStats1<-diskDataStats %>% mutate(
  CondStrain = case_when(
    condition == "shaking" & strain == "pqsl" ~ "pqsLShake",
    condition == "shaking" & strain == "wt"~ "wtShake",
    condition == "static" & strain == "pqsl"~ "pqsLStat",
    condition == "static" & strain == "wt"~ "wtStat"))
```

```{r zone of inhibition t-tests}
#the following data frames are meant to be used for t-tests
diskWtShaking<-diskData %>% filter(strain == "wt" & condition == "shaking")
diskMutShaking<-diskData %>% filter(strain == "pqsl" & condition == "shaking")
diskWtStatic<-diskData %>% filter(strain == "wt" & condition == "static")
diskMutStatic<-diskData %>% filter(strain == "pqsl" & condition == "static")

#there is no rep 1 for mutant strain, so remove first one from wt to make paired t-test
diskWtStatic1<-diskData %>% filter(strain == "wt" & condition == "static") %>% filter(!bio.rep==1)
diskWtShaking1<-diskData %>% filter(strain == "wt" & condition == "shaking")%>% filter(!bio.rep==1)

#t-tests
diskWtTtest<-t.test(x = diskWtShaking$avg, y = diskWtStatic$avg, paired = T)
diskMutTtest<-t.test(x = diskMutShaking$avg, y = diskMutStatic$avg, paired = T)
diskShakingTtest<-t.test(x = diskWtShaking1$avg, y = diskMutShaking$avg, paired = T)
diskStaticTtest<-t.test(x = diskMutStatic$avg, y = diskWtStatic1$avg, paired = T)
diskWtStaticMutShakingTtest<-t.test(x = diskMutShaking$avg, y = diskWtStatic1$avg, paired = T)
tTestData<-list(diskWtTtest, diskMutTtest,diskShakingTtest, diskStaticTtest, diskWtStaticMutShakingTtest)
TtestTable<-name <- function(dataPoint) {
  data <-data.frame(dataPoint$data.name, dataPoint$p.value)
}
tTestDataTable<-bind_rows(map(tTestData, TtestTable))
```
```{r zone of inhibition plot }
ggplot()+#both PA and Sa
  geom_errorbar(data=diskDataStats1, aes(x = strain, ymin = meanBio - SDBio, ymax = meanBio + SDBio),position = position_dodge2(width =1), width = .3)+
  geom_point(data=diskDataStats1, aes(x=strain, y=meanBio, shape =CondStrain),position = position_dodge2(width = .3), size=3)+
  scale_shape_manual(values = c(1,2,16,17),limits=c("wtStat","pqsLStat","wtShake", "pqsLShake"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type shaking")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL shaking"))))+
  scale_x_discrete(name ="", breaks = c("wt","pqsl"), labels =c("wild-type",expression(""*Delta*"pqsL")))+
  labs( y = "Zone of inhibition (mm)", shape = "Growth condtion") +
  theme_cowplot(12)

```

```{r yields}
yieldFile<-"/Users/jupabago/Documents/Whiteley/PROJECTS/pqsL SCFM2/Pa growth yields 16 hrs/Pa growth yields 16 hrs.csv"
yieldData <-read.csv(yieldFile)
yieldData %>% filter(!bio.rep==3)->yieldData
ggplot()+#both PA and Sa
  geom_point(data=yieldData, aes(x=condition, y=bugs, shape = factor(tech.rep), color = factor(bio.rep)), size=4)+
  facet_grid(.~strain)+
  scale_y_log10(limits= c(1e6,1e10))+
  labs(title =expression(paste("CFUs per mL")), x = "condtion",shape = "tech rep", color = "bio rep")

```
```{r luminescence, include = FALSE, echo=FALSE, fig.width=12, fig.height=9, results='asis', warning=FALSE, message=FALSE}
luminescenceFile<-"/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Staph killing assays/summary.csv"
luminescenceData <-read.csv(luminescenceFile)
luminescenceData<-luminescenceData %>% group_by(TimePoint, BioReplicate, Condition, Strain) %>% mutate(techMean = mean(Luminescence)) %>% ungroup() %>% group_by(TimePoint, Condition, Strain) %>% mutate(Mean = mean(techMean), SE =std.error(techMean)) %>% ungroup()
#luminescenceData1<-luminescenceData %>%  
#mutate(LogTotals = ifelse(Totals>0, log10(Totals), 0))#this adds up all the data on each slice and then gets the log of that total

luminescenceData1<-luminescenceData %>% mutate(
  CondStrain = case_when(
    Condition == "shaking" & Strain == "pqsL" ~ "pqsLShake",
    Condition == "shaking" & Strain == "wt"~ "wtShake",
    Condition == "static" & Strain == "pqsL"~ "pqsLStat",
    Condition == "static" & Strain == "wt"~ "wtStat"))

ggplot()+#both PA and Sa
  #geom_point(data=luminescenceData, aes(x=TimePoint, y=techMean, shape = Strain, color = Condition), size=1)+
  geom_line(data=luminescenceData1, aes(x=TimePoint, y=Mean, linetype = Condition, group = interaction (Strain, Condition)), size=.5,show.legend = FALSE)+
  geom_point(data=luminescenceData1, aes(x=TimePoint, y=Mean, shape = CondStrain), size=3)+
  scale_y_log10(name=expression(paste(italic('S. aureus'), " % luminescence vs control" )), breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs( x = "Time (minutes)",shape = "") +
  scale_shape_manual(values = c(1,2,16,17),limits=c("wtStat","pqsLStat","wtShake", "pqsLShake"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type shaking")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL shaking"))))+
  scale_linetype_manual(values =c("solid","dashed"))+
  theme_cowplot(12)

```

```{r Staph TOB MIC}
micFile<-"/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/SA TOB MIC/SA TOB MIC.csv"
micData <-read.csv(micFile)
micData<-micData %>% filter(!concentration==0.5) %>% mutate(LogCounts = log10(value)) %>% group_by(concentration, condition) %>% mutate(meanLog = mean(LogCounts), seLog = std.error(LogCounts)) %>% ungroup() %>% group_by(replicate, condition)%>% mutate(LogDifference = LogCounts[concentration==0]-LogCounts) %>% ungroup() %>% group_by(concentration, condition) %>% mutate(meanLogDifference = mean(LogDifference), seLogDifference = std.error(LogDifference)) %>% filter(!concentration==0)
controls<-micData %>% filter(concentration==0)
cleanMicData<-anti_join(micData,controls)

ggplot()+
  geom_point(data=cleanMicData, aes(x=concentration, y=LogCounts, color = condition), size=2)+
  geom_line(data=cleanMicData, aes(x=concentration, y=meanLog, color = condition), size=2)+
  scale_x_log10(name=expression(paste("[TOB] ug/mL" )), breaks = c(2,8,32,64,128,256,512), labels =c(2,8,32,64,128,256,512))+
  scale_y_continuous(name = "Log CFUs", breaks = c(0,2,4,6,8,10), labels =c(0,2,4,6,8,10))+
  labs(title =expression(paste("S. aureus TOB MIC")), x = "[TOB] ug/mL",shape = "tech rep", color = "Condition")+
  scale_color_grey()+
  theme_cowplot(12)

ggplot()+
  geom_col(data=cleanMicData, aes(x=as.factor(concentration), y=meanLogDifference, fill = condition ),position =position_dodge(width = .5), width = .5)+
  geom_linerange(data = cleanMicData, aes(x = as.factor(concentration), ymin = meanLogDifference - seLogDifference, ymax = meanLogDifference + seLogDifference, group = condition),position = position_dodge(width = .5))+
  scale_fill_grey()+
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.05)), breaks = c(2,4,6,8,10), labels = c(2,4,6,8,10))+
  labs(title =expression(paste("S. aureus TOB MIC")), x = "[TOB] ug/mL",y = "log CFU Change", color = "Condition")+
  theme_cowplot()+
  theme_half_open()+
  background_grid()

  geom_line(data=cleanMicData, aes(x=concentration, y=meanLog, color = condition), size=2)+

  scale_y_continuous(name = "Log CFUs", breaks = c(0,2,4,6,8,10), labels =c(0,2,4,6,8,10))+
  labs(title =expression(paste("S. aureus TOB MIC")), x = "[TOB] ug/mL",shape = "tech rep", color = "Condition")+
  theme_cowplot(12)
```

```{r final plots, version 1}
#this is the original Figure but it needs to be split in two parts
    DiskPlot<-ggplot()+
      geom_errorbar(data=diskDataStats1, aes(x = strain, ymin = meanBio - SDBio, ymax = meanBio + SDBio, group = condition),position = position_dodge(width = .6), width = .2)+
      geom_point(data=diskDataStats1, aes(x=strain, y=meanBio, shape =CondStrain),position = position_dodge2(width = .6), size=3)+
      scale_shape_manual(values = c(1,2,16,17),limits=c("wtStat","pqsLStat","wtShake", "pqsLShake"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type stirred")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL stirred"))))+
      scale_x_discrete(name ="", breaks = c("wt","pqsl"), labels =c("wild-type",expression(""*Delta*"pqsL")))+
      labs( y = "Zone of inhibition (mm)", shape = "Growth condtion") +
      theme_cowplot(12)

    luminescencePlot<-ggplot()+
      geom_line(data=luminescenceData1, aes(x=TimePoint, y=Mean, linetype = Condition, group = interaction (Strain, Condition)), size=.5,show.legend = FALSE)+
      geom_point(data=luminescenceData1, aes(x=TimePoint, y=Mean, shape = CondStrain), size=3)+
      scale_y_log10(name=expression(paste(italic('S. aureus'), " % luminescence vs control" )))+
      labs( x = "Time (minutes)",shape = "") +
      scale_shape_manual(values = c(1,2,16,17),limits=c("wtStat","pqsLStat","wtShake", "pqsLShake"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type stirred")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL stirred"))))+
      scale_linetype_manual(values =c("solid","dashed"))+
      theme(legend.position="top")+
      theme_cowplot(12)

    legendFigure2 <- get_legend(luminescencePlot + theme(legend.box.margin = margin(0, 0, 0, -10)))
    figure2 <- plot_grid(DiskPlot+theme(legend.position="none"),luminescencePlot+theme(legend.position="none"),
                      align='vh',labels = c("A", "B"),hjust = -1,nrow = 1)
    plot_grid(figure2,legendFigure2,rel_widths = c(6, 1.7))
```
```{r final plots, version 2}
#this figure needs to be split to show only wt data and mutant data and limit the time point to 7 minutes for the luminescence.

    DiskPlotWt<-ggplot()+
      geom_errorbar(data=diskDataStats1 %>% filter (strain == "wt"), aes(x = condition, ymin = meanBio - SDBio, ymax = meanBio + SDBio),width = .2)+
      geom_point(data=diskDataStats1 %>% filter (strain == "wt"), aes(x=condition, y=meanBio, shape =condition), size=3)+
      scale_shape_manual(values = c(1,2),limits=c("static","shaking"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type stirred"))))+
      scale_x_discrete(name ="", breaks = c("static","shaking"), labels =c("Static","Stirred"))+
    ylim(0,17.5)+
  labs( y = "Zone of inhibition (mm)", shape = "Growth condtion") +
      theme_cowplot(12)

    DiskPlotMut<-ggplot()+
      geom_errorbar(data=diskDataStats1 %>% filter (strain == "pqsl"), aes(x = condition, ymin = meanBio - SDBio, ymax = meanBio + SDBio),width = .2)+
      geom_point(data=diskDataStats1 %>% filter (strain == "pqsl"), aes(x=condition, y=meanBio, shape =condition), size=3)+
      scale_shape_manual(values = c(16,17),limits=c("static","shaking"), 
                     labels = c(bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL stirred"))))+
      ylim(0,17.5)+
      scale_x_discrete(name ="", breaks = c("static","shaking"), labels =c("Static","Stirred"))+
      labs( y = "Zone of inhibition (mm)", shape = "Growth condtion") +
      theme_cowplot(12)

    luminescencePlotWt<-ggplot()+
      geom_line(data=luminescenceData1 %>% filter (Strain == "wt") %>% filter (TimePoint<7), aes(x=TimePoint, y=Mean, group = Condition), size=.5, show.legend = FALSE)+
      geom_point(data=luminescenceData1%>% filter (Strain == "wt") %>% filter (TimePoint<7), aes(x=TimePoint, y=Mean, shape = Condition), size=3)+
      scale_y_log10(name=expression(paste(italic('S. aureus'), " % luminescence vs control" )), limits = c(1,121),labels = c(1, 10, 100),breaks = c(1, 10, 100))+
      labs( x = "Time (minutes)",shape = "") +
      scale_shape_manual(values = c(1,2),limits=c("static","shaking"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type stirred"))))+
      theme(legend.position="top")+
      theme_cowplot(12)

    luminescencePlotMut<-
      ggplot()+
      geom_line(data=luminescenceData1 %>% filter (Strain == "pqsL") %>% filter (TimePoint<8), aes(x=TimePoint, y=Mean, group = Condition), size=.5, show.legend = FALSE)+
      geom_point(data=luminescenceData1%>% filter (Strain == "pqsL") %>% filter (TimePoint<8), aes(x=TimePoint, y=Mean, shape = Condition), size=3)+
      scale_y_log10(name=expression(paste(italic('S. aureus'), " % luminescence vs control" )),limits = c(1,121), labels = c(1,10,100),breaks = c(1, 10, 100))+
      labs( x = "Time (minutes)",shape = "") +
      scale_shape_manual(values = c(16,17),limits=c("static", "shaking"), 
                     labels = c(bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL stirred"))))+
      theme(legend.position="top")+
      theme_cowplot(12)
#Figure 2

    legendFigure2 <- get_legend(DiskPlotWt + theme(legend.box.margin = margin(0, 0, 0, -10)))
    figure2 <- plot_grid(DiskPlotWt+theme(legend.position="none"),luminescencePlotWt+theme(legend.position="none"),
                      align='vh',labels = c("A", "B"),hjust = -1,nrow = 1)
    plot_grid(figure2,legendFigure2,rel_widths = c(6, 1.7))

#Figure 5
    legendFigure5 <- get_legend(DiskPlotMut + theme(legend.box.margin = margin(0, 0, 0, -10)))
    figure5 <- plot_grid(DiskPlotMut+theme(legend.position="none"),luminescencePlotMut+theme(legend.position="none"),
                      align='vh',labels = c("A", "B"),hjust = -1,nrow = 1)
    plot_grid(figure5,legendFigure5,rel_widths = c(6, 1.7))

```